### 2025-03-02 article.tex
innitial figure done

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath, amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\usepackage{pgfplots}
\pgfplotsset{compat=1.17}

\title{Comparing the Efficiency of Machine Learning and Deep Learning Models in Source Code Authorship Attribution}
\author{Richard Čerňanský}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
\end{abstract}

\section{Introduction}
Introduction

\section{Methods}
Methods

\section{Results}

\begin{table}[h!]
    \centering
    {\scriptsize
    \begin{tabular}{lcccc}
    \hline
    \multicolumn{5}{c}{\textbf{Models' Author Attribution Accuracy for Different Author Set Sizes}} \\
    \hline
    \textbf{Model} & \multicolumn{4}{c}{\textbf{Author Set Size}} \\
    \cline{2-5}
                  & 110  & 27  & 11 & 3 \\ 
    \hline
    Random Forests (TF--IDFs w/o comments) & 66.24\%\;|\;$3.63\,\mathrm{min}$ & 70.60\%\;|\;$0.23\,\mathrm{min}$ & 80.56\%\;|\;$0.08\,\mathrm{min}$ & 85.84\%\;|\;$0.05\,\mathrm{min}$ \\
    BERT (source code w/o comments)        & 72.47\%\;|\;$70.33\,\mathrm{min}$ & 82.72\%\;|\;$32.49\,\mathrm{min}$ & 89.84\%\;|\;$9.34\,\mathrm{min}$  & 90.70\%\;|\;$1.82\,\mathrm{min}$ \\
    BERT (AST pre\_order traversal)        & 24.07\%\;|\;$73.53\,\mathrm{min}$ & 35.66\%\;|\;$30.00\,\mathrm{min}$ & 40.62\%\;|\;$5.69\,\mathrm{min}$  & 83.72\%\;|\;$1.22\,\mathrm{min}$ \\
    AttentionNN (AST paths)              & --- & --- & --- & --- \\
    \hline
    \end{tabular}
    }
    \caption{Comparison of models across author attribution accuracy and training time for different author set sizes.}
    \label{tab:comparison}
\end{table}

\begin{figure}[h!]
\centering
\begin{tikzpicture}
\begin{axis}[
    view={60}{30},             % 3D viewing angles (azimuth, elevation)
    z buffer=sort,             % Ensures correct front-to-back rendering
    xlabel={Model},
    xlabel style={
        xshift = -40pt,
        yshift=40pt,
        align=center        % optional alignment
    },
    ylabel={Author set size},
    ylabel style={
        xshift = 25pt,
        yshift=20pt,
        align=center        % optional alignment
    },
    zlabel={Metric (\%)},
    width=14cm,
    height=8cm,
    xtick={1,2,3,4},
    xticklabels={RF (TF-IDF), BERT (source code), BERT (AST pre\_order), AttentionNN},
    xticklabel style={
        font=\footnotesize,   % or \footnotesize, \tiny, etc.
        xshift = -20pt,
        yshift=10pt,
        align=center        % optional alignment
    },
    ytick={1,2,3,4},
    yticklabels={3 (min. count 58), 11 (min. count 50), 27 (min. count 40), 110 (min. count 25)},
    yticklabel style={
        font=\footnotesize,   % or \footnotesize, \tiny, etc.
        xshift = 25pt,
        yshift=5pt,
        align=center        % optional alignment
    },
    zmin=0, zmax=100,
    grid=major,
    legend style={
        at={(0.5,1.15)},
        anchor=north,
        legend columns=3
    }
]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MODEL 1: RF (x = 1)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\addplot3+[ybar, bar width=12pt, fill=blue, draw=none, bar shift=-0.2] 
    coordinates {(1.1,1,85.84) (1,2,80.56) (1,3,70.60) (1,4,66.24)};
\addplot3+[ybar, bar width=12pt, fill=red, draw=none, bar shift=0]  
    coordinates {(1.1,1,89.36) (1,2,84.56) (1,3,79.27) (1,4,71.73)};
\addplot3+[ybar, bar width=12pt, fill=brown, draw=none, bar shift=+0.2] 
    coordinates {(1.1,1,84.96) (1,2,72.03) (1,3,69.73) (1,4,65.55)};

\addplot3+[
    mark=none,
    blue,
    dotted,
    line width=1pt,
    nodes near coords,
    every node near coord/.append style={
        mark=none,
        fill=none,
        draw=none,
        text=black,
        font=\scriptsize,
        yshift=0pt
    }
] coordinates {
    (0.8,1,85.84) (0.8,2,80.56) (0.8,3,70.60) (0.8,4,66.24)
};

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MODEL 2: BERT-SOURCE (x = 2)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\addplot3+[ybar, bar width=12pt, fill=blue, draw=none, bar shift=-0.2] 
    coordinates {(2,1,90.70) (2,2,89.84) (2,3,82.72) (2,4,72.47)};
\addplot3+[ybar, bar width=12pt, fill=red, draw=none, bar shift=0]  
    coordinates {(2,1,93.33) (2,2,90.33) (2,3,84.66) (2,4,76.69)};
\addplot3+[ybar, bar width=12pt, fill=brown, draw=none, bar shift=+0.2] 
    coordinates {(2,1,91.67) (2,2,89.92) (2,3,82.25) (2,4,71.76)};

\addplot3+[
    mark=none,
    blue,
    dotted,
    line width=1pt,
    nodes near coords,
    every node near coord/.append style={
        mark=none,
        fill=none,
        draw=none,
        text=black,
        font=\scriptsize,
        yshift=0pt
    }
] coordinates {
    (1.8,1,90.70) (1.8,2,89.84) (1.8,3,82.72) (1.8,4,72.47)
};

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MODEL 3: BERT-AST (x = 3)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\addplot3+[ybar, bar width=12pt, fill=blue, draw=none, bar shift=-0.2] 
    coordinates {(3,1,83.72) (3,2,40.62) (3,3,35.66) (3,4,24.07)};
\addplot3+[ybar, bar width=12pt, fill=red, draw=none, bar shift=0]  
    coordinates {(3,1,84.72) (3,2,40.53) (3,3,36.41) (3,4,27.40)};
\addplot3+[ybar, bar width=12pt, fill=brown, draw=none, bar shift=+0.2] 
    coordinates {(3,1,82.58) (3,2,38.30) (3,3,33.93) (3,4,23.33)};

\addplot3+[
    mark=none,
    blue,
    dotted,
    line width=1pt,
    nodes near coords,
    every node near coord/.append style={
        mark=none,
        fill=none,
        draw=none,
        text=black,
        font=\scriptsize,
        yshift=0pt
    }
] coordinates {
    (2.8,1,83.72) (2.8,2,40.62) (2.8,3,35.66) (2.8,4,24.07)
};

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MODEL 4: ATTENTION-NN (x = 4) - Data missing => set to 0
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\foreach \yy in {1,2,3,4} {
    \addplot3+[ybar, bar width=12pt, fill=blue, draw=none, bar shift=-0.2] coordinates {(4,\yy,0)};
    \addplot3+[ybar, bar width=12pt, fill=red, draw=none, bar shift=0] coordinates {(4,\yy,0)};
    \addplot3+[ybar, bar width=12pt, fill=brown, draw=none, bar shift=+0.2] coordinates {(4,\yy,0)};
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% LEGEND
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\addplot3[
    ybar, 
    fill=blue, 
    draw=none, 
    bar width=12pt, 
    legend image code/.code={
        \draw[fill=blue] (0cm,-0.1cm) rectangle (0.6cm,0.1cm);
    }
] coordinates {(0,0,-9999)};
\addlegendentry{Accuracy}

\addplot3[
    ybar, 
    fill=red, 
    draw=none, 
    bar width=12pt, 
    legend image code/.code={
        \draw[fill=red] (0cm,-0.1cm) rectangle (0.6cm,0.1cm);
    }
] coordinates {(0,0,-9999)};
\addlegendentry{Precision}

\addplot3[ % Recall
    ybar, 
    fill=brown, 
    draw=none, 
    bar width=12pt, 
    legend image code/.code={
        \draw[brown, mark=x] plot coordinates {(0cm,0cm)}; 
    }
] coordinates {(0,0,-9999)};
\addlegendentry{Recall}

\end{axis}
\end{tikzpicture}
\caption{Bar chart comparing Accuracy (with accuracy values labeled), Precision, and Recall across models and author set sizes with minimum function count per author. }
\label{fig:3d-bars-legend-brown}
\end{figure}

\section{Discussion}
Here we discuss the implications of our results.

\begin{figure}[h]
\centering
\includegraphics[width=0.5\textwidth]{example-image} % Uses the default dummy image provided by the graphicx package
\caption{A sample figure using the default example image.}
\end{figure}

\section{Conclusion}
We conclude that this document compiles successfully. If you can read all sections, view the figure, and see the mathematical expressions, your LaTeX environment is correctly set up.

\end{document}

### 2025-03-02 article.tex
figure in progress... but almost done

@@
     },
-    ytick={1,2,3,4},
-    yticklabels={3 (min. count 58), 11 (min. count 50), 27 (min. count 40), 110 (min. count 25)},
+    ytick={4,3,2,1}, % Reversed ytick values
+    yticklabels={3 (min. count 58), 11 (min. count 50), 27 (min. count 40), 110 (min. count 25)}, % Corresponding reversed labels
     yticklabel style={
@@
 \addplot3+[ybar, bar width=12pt, fill=blue, draw=none, bar shift=-0.2]
-    coordinates {(1.1,1,85.84) (1,2,80.56) (1,3,70.60) (1,4,66.24)};
+    coordinates {(1.1,4,85.84) (1,3,80.56) (1,2,70.60) (1,1,66.24)}; % Coordinates are adjusted to match the reversed ytick
 \addplot3+[ybar, bar width=12pt, fill=red, draw=none, bar shift=0]
-    coordinates {(1.1,1,89.36) (1,2,84.56) (1,3,79.27) (1,4,71.73)};
+    coordinates {(1.1,4,89.36) (1,3,84.56) (1,2,79.27) (1,1,71.73)}; % Coordinates are adjusted to match the reversed ytick
 \addplot3+[ybar, bar width=12pt, fill=brown, draw=none, bar shift=+0.2]
-    coordinates {(1.1,1,84.96) (1,2,72.03) (1,3,69.73) (1,4,65.55)};
+    coordinates {(1.1,4,84.96) (1,3,72.03) (1,2,69.73) (1,1,65.55)}; % Coordinates are adjusted to match the reversed ytick
 
@@
 ] coordinates {
-    (0.8,1,85.84) (0.8,2,80.56) (0.8,3,70.60) (0.8,4,66.24)
+    (0.8,4,85.84) (0.8,3,80.56) (0.8,2,70.60) (0.8,1,66.24) % Coordinates are adjusted to match the reversed ytick
 };
@@
 \addplot3+[ybar, bar width=12pt, fill=blue, draw=none, bar shift=-0.2]
-    coordinates {(2,1,90.70) (2,2,89.84) (2,3,82.72) (2,4,72.47)};
+    coordinates {(2,4,90.70) (2,3,89.84) (2,2,82.72) (2,1,72.47)}; % Coordinates are adjusted to match the reversed ytick
 \addplot3+[ybar, bar width=12pt, fill=red, draw=none, bar shift=0]
-    coordinates {(2,1,93.33) (2,2,90.33) (2,3,84.66) (2,4,76.69)};
+    coordinates {(2,4,93.33) (2,3,90.33) (2,2,84.66) (2,1,76.69)}; % Coordinates are adjusted to match the reversed ytick
 \addplot3+[ybar, bar width=12pt, fill=brown, draw=none, bar shift=+0.2]
-    coordinates {(2,1,91.67) (2,2,89.92) (2,3,82.25) (2,4,71.76)};
+    coordinates {(2,4,91.67) (2,3,89.92) (2,2,82.25) (2,1,71.76)}; % Coordinates are adjusted to match the reversed ytick
 
@@
 ] coordinates {
-    (1.8,1,90.70) (1.8,2,89.84) (1.8,3,82.72) (1.8,4,72.47)
+    (1.8,4,90.70) (1.8,3,89.84) (1.8,2,82.72) (1.8,1,72.47) % Coordinates are adjusted to match the reversed ytick
 };
@@
 \addplot3+[ybar, bar width=12pt, fill=blue, draw=none, bar shift=-0.2]
-    coordinates {(3,1,83.72) (3,2,40.62) (3,3,35.66) (3,4,24.07)};
+    coordinates {(3,4,83.72) (3,3,40.62) (3,2,35.66) (3,1,24.07)}; % Coordinates are adjusted to match the reversed ytick
 \addplot3+[ybar, bar width=12pt, fill=red, draw=none, bar shift=0]
-    coordinates {(3,1,84.72) (3,2,40.53) (3,3,36.41) (3,4,27.40)};
+    coordinates {(3,4,84.72) (3,3,40.53) (3,2,36.41) (3,1,27.40)}; % Coordinates are adjusted to match the reversed ytick
 \addplot3+[ybar, bar width=12pt, fill=brown, draw=none, bar shift=+0.2]
-    coordinates {(3,1,82.58) (3,2,38.30) (3,3,33.93) (3,4,23.33)};
+    coordinates {(3,4,82.58) (3,3,38.30) (3,2,33.93) (3,1,23.33)}; % Coordinates are adjusted to match the reversed ytick
 
@@
 ] coordinates {
-    (2.8,1,83.72) (2.8,2,40.62) (2.8,3,35.66) (2.8,4,24.07)
+    (2.8,4,83.72) (2.8,3,40.62) (2.8,2,35.66) (2.8,1,24.07) % Coordinates are adjusted to match the reversed ytick
 };

### 2025-03-02 article.tex
fixed bug of dottted line

@@
 
+    %dotted line
 \addplot3+[
@@
     coordinates {(3,4,82.58) (3,3,38.30) (3,2,33.93) (3,1,23.33)}; % Coordinates are adjusted to match the reversed ytick
-
+\foreach \yA/\zA/\yB/\zB in {
+    4/83.72/3/40.62,
+    3/40.62/2/35.66,
+    2/35.66/1/24.07
+} {
     \addplot3+[
@@
     ] coordinates {
-    (2.8,4,83.72) (2.8,3,40.62) (2.8,2,35.66) (2.8,1,24.07) % Coordinates are adjusted to match the reversed ytick
+        (3,\yA,\zA) (3,\yB,\zB)
     };
+}
 
@@
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.5\textwidth]{example-image} % Uses the default dummy image provided by the graphicx package
-\caption{A sample figure using the default example image.}
-\end{figure}
-
 \section{Conclusion}

### 2025-03-10 article.tex
article missing some refs

@@
 \usepackage{graphicx}
+\usepackage{float} 
 \usepackage{hyperref}
+\usepackage{subcaption}
 \usepackage{geometry}
+\usepackage{bookmark}
 \geometry{margin=1in}
-
 \usepackage{pgfplots}
@@
 
-\title{Comparing the Efficiency of Machine Learning and Deep Learning Models in Source Code Authorship Attribution}
-\author{Richard Čerňanský}
+\title{Comparing Machine Learning and Deep Learning Models with Attention Mechanisms for Multiple Authorship Attribution}
+\author{Richard Čerňanský, Ing. Juraj Petrík}
 \date{\today}
@@
 \begin{abstract}
+    Authorship attribution (AA) in source code is crucial for plagiarism detection, 
+    forensic analysis, and software maintenance, especially when author 
+    information is missing or ambiguous. This field is not as well explored 
+    as the written communication AA and there are new models to be tested 
+    on different sized dataset. This study compares traditional 
+    machine learning models, such as Random Forests trained on TF-IDF 
+    tokenized representations, with deep learning models, including BERT 
+    and an Attention-based Neural Network (AttentionNN), for the task of 
+    multiple authorship attribution. We evaluate different representations 
+    of source code, including raw tokenized text and Abstract Syntax Tree (AST) 
+    structures, to determine their effectiveness in distinguishing authorship. 
+    The models are tested on a dataset of 3,004 C language functions from 
+    Google Code Jam spanning 2009–2018, assessing their accuracy, precision, 
+    recall, and training efficiency across varying author set sizes. 
+    Our findings highlight the strengths and limitations of each approach, 
+    showing that while deep learning models, particularly BERT trained on 
+    raw source code, achieve the highest accuracy, Random Forests offer 
+    competitive results with significantly lower computational cost. 
+    The study also explores the impact of function complexity on model 
+    performance, revealing that AttentionNN benefits from structural 
+    complexity while struggling with short functions. The results emphasize 
+    the trade-offs between accuracy, computational efficiency, and data 
+    representation, providing insights into the optimal model selection 
+    for source code authorship attribution.   
 \end{abstract}
@@
 \section{Introduction}
-Introduction
+
+Authorship attribution in source code analysis is a crucial task 
+with applications in plagiarism detection, forensic investigations, and 
+intellectual property protection, but also software maintenance
+and software quality analysis, where the information about an author 
+is missing, inaccurate or ambiguous.
+Given a function written in a programming
+language, determining its author requires capturing distinct stylistic 
+patterns. Traditional machine learning (ML) methods, such as Random 
+Forests trained on TF-IDF tokenized representations, have shown promise 
+in text-based authorship identification. However, recent advancements 
+in deep learning (DL) have introduced transformer-based models, such as 
+BERT, which leverage contextual embeddings and self-attention mechanisms 
+to extract more meaningful patterns from textual data.
+
+Despite the success of deep learning models in natural language 
+processing (NLP), their application to source code authorship attribution on different dataset sizes. 
+remains relatively unexplored. Unlike natural language, source code 
+exhibits a more rigid syntax and structure, making its representation 
+a key challenge. Several approaches have been proposed, ranging from 
+direct tokenized representations of the raw source code to abstract 
+syntax tree (AST) transformations that capture program logic. However, 
+it remains unclear which type of representation is most effective for 
+distinguishing authorship across different levels of dataset complexity.
+
+This study systematically compares machine learning and deep learning 
+models, including Random Forests, BERT trained on tokenized source code, 
+BERT trained on AST representations, and an Attention-based Neural Network 
+(AttentionNN) utilizing AST-derived node-to-node paths. The evaluation is 
+conducted on a dataset of 3,004 C language functions from Google Code Jam 
+(GCJ) spanning 2009 to 2018 \cite{gcj}. The models are assessed based on their 
+accuracy, precision, recall, and training efficiency across different 
+author set sizes.
+
+Through this comparison, we aim to determine which model best balances 
+accuracy and computational efficiency, as well as the effectiveness of 
+different function representations for authorship attribution.
 
 \section{Methods}
-Methods
+
+\subsection{Source code representation}
+
+In the paper \cite{information}, authors present division of features based on the method they use for extraction:
+
+\begin{itemize}
+
+    \item \textbf{Stylometric features:} 
+    They express the structural characters that represent the author’s preference for different statements, keywords, nesting sizes, etc.
+    \item \textbf{N-gram features:}
+    N-gram is a contiguous sequence of an n byte, character, or words extracted from the source. A sliding that has a fixed length generates 
+    the sequence.
+    \item \textbf{Graph based features:}
+    These capture underlying traits in the code such as deeply nesting code, control flow and data flow.
+    It includes : AST, Program Dependency Graph (PDG), Control Flow Graph (CFG), RFG.
+    \item \textbf{Behavioral features:}
+    This type completely different from the previous as it is not looking at the source code itself
+    but rather on the way that the program of the source code communicates with the computer.
+    In this category there are dynamic features which collect runtime measurements as CPU and memory usage.
+    Instruction features that are retrieved from binaries and reflect code properties like registers or immediate operands.
+    \item \textbf{Neural Networks Generated Representations of Source Code or Features:}
+    Lastly, in recent years, neural networks have been increasingly utilized 
+    to generate representations (such as embeddings) of source code. This 
+    approach is also applied in this thesis, using AST analysis to provide 
+    input for the neural network.
+
+\end{itemize}
+
+\subsection{Abstract Syntax Tree}
+
+\quad Based on the full definition \cite{sun2023AST} Abstract syntax tree (AST) is a tree-based representation of source code, where nodes represent various elements of the source code 
+and paths represent relationships of the structure. AST is an inseparable part of compilation process for many reasons. It is used as an 
+intermediate representation of program utilized for optimization and generation of machine code. Nodes of the tree can be either internal 
+(non-leaf) or leafs. Internal nodes define the program's constructs or operations for their children. Leaf nodes store the actual textual value.
+
+Below is an example of a simple C function and its corresponding Abstract Syntax Tree (AST):
+\begin{verbatim}
+int add(int a, int b) {
+    return a + b;
+}
+\end{verbatim}
+
+The corresponding AST is structured as follows:
+\begin{verbatim}
+TranslationUnit
+|--FunctionDefinition `int add(int a, int b) {`
+   |--BasicTypeSpecifier `int`
+   |--FunctionDeclarator `add(`
+   |  |--IdentifierDeclarator `add`
+   |  |--ParameterSuffix `(int a, int b)`
+   |     |--ParameterDeclaration `int a`
+   |     |  |--BasicTypeSpecifier `int`
+   |     |  |--IdentifierDeclarator `a`
+   |     |--ParameterDeclaration `int b`
+   |        |--BasicTypeSpecifier `int`
+   |        |--IdentifierDeclarator `b`
+   |--CompoundStatement `{ return a + b; }`
+      |--ReturnStatement `return a + b;`
+         |--AddExpression `a + b`
+            |--IdentifierName `a`
+            |--IdentifierName `b`
+\end{verbatim}
+\label{verbatim:AST_example}
+
+\subsubsection{AST - types of features}
+
+Features divide into 4 main kinds according to the nature of the information extracted from the Abstract Syntax Tree (AST):
+\begin{itemize}
+    \item \textbf{Structural} \label{item:structural}
+    Structural features capture the structural complexity of AST.
+These are often some numerical quantitative values like depth of the graph, number of nodes, or average branching factor of internal nodes.
+    \item \textbf{Semantic} \\
+    Semantic features reflect the semantic information encoded in AST.
+It could be as simple as the distribution of node kinds (that are defined by the compiler that creates the AST) or the distribution of distinct root-to-leaf paths.
+    \item \textbf{Syntactic} \\
+    Syntactic features describe the paradigm in which the source code is written and the logical complexity of the program.
+These might include the number of functions or functor structures, or the number of control flow units.
+    \item \textbf{Combined} \\
+    Combined features are use-case specific, and it is up to the data analyst to design the best fit solution for information extraction using the combinations and alternations of the aforementioned strategies.
+Combined features are also the ones that are usually used when tackling real-world problems like source code authorship attribution or function name classification.
+\end{itemize}
+
+\subsubsection{Halstead complexity measures}
+
+For evaluation of complexity of examined functions we have used Halstead complexity 
+measures Difficulty, Volume, Effort \cite{halstead}. To aggregate them into single value, 
+we used weighted sum of the three metrics weighting them in ratios 4:3:3 respectively. 
+We did not incorporate Halstead length because we did separate analysis on length.
+
+\subsection{Our Data Representation}
+For training the models, we constructed our prediction task data pipeline like this:
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=2.5cm, height=3.5cm]{figures/high_level_prediciton_task.png}
+    \caption{\cite{compiler_based} High level overview of the function classification preditction task.}
+    \label{fig:predictive_task}
+\end{figure}
+
+We utilized various 
+types of representations of single function's raw source code to provide the models with data.
+These include word tokenized TF-IDF vectors for Random Forests, BPE of raw source code for
+BERT, linearized AST pre-order traversal for BERT and representation set of randomly selected 600 node-to-node paths for AttentionNN.  
+PsycheC project's C language frontend parser \cite{psychec} was used for parsing the code into ASTs.
+Each model was trained in Tensorflow.Keras framework and with different author set sizes to evaluate scalabillity.
+The different set sizes were
+110, 27, 11, 3 with each having increased demands on minimal function count than the one before. The minimal function counts per author for the sizes were 25, 40, 50, 58 respectively.
+
+\subsection{Data preprocessing \& Exploratory data analysis}
+The dataset consisted of thousands of solutions from hundreds of authors. From each solution we extracted the function and its author.
+The largest of the different sized datasets we used consisted of $3\,004$ C language functions' source code gathered from Google Code Jam (GCJ) solutions from years 2009 to 2018.
+The data was preprocessed exclude duplicates,
+remove outlying authors that have multiple times more samples than is the average number of samples per author to reduce the models' bias towards these names. Functions 
+that could not be parsed into an AST were also eliminated to ensure consistency in the datasets. This allows for a fair assessment of each model, 
+as they utilize different representations of function data. For further analysis, we needed to see the frequency distributions of authors.
+
+\begin{figure}[h]
+    \centering
+    \begin{subfigure}{0.48\textwidth} % Adjust width to fit
+        \centering
+        \includegraphics[width=\linewidth]{figures/distribution_names.pdf}
+        \caption{Frequency distribution of author names}
+        \label{fig:distr_names}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.48\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{figures/distr_lens_tokens.pdf}
+        \caption{Frequency distribution of function lengths}
+        \label{fig:disrt_lens}
+    \end{subfigure}
+    \caption{Important frequency distributions of our dataset}
+    \label{fig:eda_distr}
+\end{figure}
+
+From the figure \ref{fig:distr_names} we can see that the most of our dataset's 
+functions lengths lie between 10 and 50 tokens which makes our goal of authorship attribution of single function
+harder than if given a whole file consisting of multiple functions or code blocks in sense that for the prediction
+most of the time we only have from 10 to 50 tokens available to find the correct author.
+
+\section{Models setup \& training}
+
+In this section we will discuss the training strategies and setup for the examined models.
+Random Forests and AttentionNN models were trained using Stratified Cross validation strategy that ensures, that each author is evenly distributed
+across the folds. Number of folds was set to 5 and 3 folds were used for training, 1 for validation and 1 for testing.
+As training of BERT takes a lot more time than the other 2, we just repeated the traing 3 times and 
+observed variance in evaluation metrics lower than 1\%.
+
+\subsection{Random Forests}
+
+For comparison purposes, we have chosen Random Forests as standard approach to text based authorship attribution. 
+They were trained on word-level tokens using the regex pattern \verb|r"\b\w+\b"|, which matches words separated by word boundary characters.
+Firt we used grid search to optimize these hyperparameters:
+\begin{itemize}
+    \item \textbf{n\_estimators}: \{50, 100, 200, 500\}
+    \item \textbf{max\_depth}: \{10, 20, 30, None\}
+    \item \textbf{min\_samples\_split}: \{2, 5, 10\}
+    \item \textbf{min\_samples\_leaf}: \{1, 2, 4\}
+    \item \textbf{max\_features}: \{"sqrt", "log2", None\}
+    \item \textbf{bootstrap}: \{True, False\}
+\end{itemize}
+
+and then we trained the model using the best hyperparameters found.
+
+\subsection{BERT}
+
+For both BERT training setups we used Microsoft's pretrained graphcodebert-base \cite{codebert} that already understands general programming concepts like 
+control structures, loops or functions and is able to capture differences between authors faster. We also used the graphcodebert tokenizer which uses 
+Byte pair encoding that merges frequently occurring character sequences into tokens, which is appropriate for data like source code of programming language
+that has very rigid syntax. The comments in all code samples were filtered out so the attribution task
+is 'fair' across the models that utilize ASTs.
+
+As mentioned, we have trained the BERT model on two distinct data representations. Tokenized RAW source code and tokenized AST pre-order traversal
+to see how the model reacts to these representations and how useful it may be to learn from the AST structural patterns in comparison to AttentionNN 
+that utilizes node-to-node paths.
+
+\subsection{Attention Neural Network}
+
+To capture all three of the structural, semantic and syntactic features in a number vector that could be somehow fed into a neural network classifier. 
+Code2vec \cite{code2vec} proposes solution design that extracts so called 'path-contexts' which are encoded into vector space using the embedding values 
+of its components.
+Path-contexts capture the node-to-node path in concise manner as it consist of triplets of <start node, the path of nodes kinds, terminal node>.
+Each distinct node is encoded with the data value it carries and each distinct path is encoded as a string of nodes' kinds in between the start and terminal node.
+The model then learns to distinguish the authors utilizing the attention mechanism, giving larger weights to contexts that help to reveil the author the most.
+
+\subsection{Evaluation methods}
+For evaluation purposes metrics such as validation Accuracy, Precision, Recall, Time to train model and Halstead complexity measures were captured to explain the models
+behavior with different kinds of data. The deep-learning models were trained with the Categorical Cross-Entropy loss function guiding the optimization process,
+
+\begin{equation}
+    \mathcal{L}_i = - \sum_{j=1}^{C} y_{i,j} \log(\hat{y}_{i,j})
+\end{equation}
+
+where:
+\begin{itemize}
+    \item $\mathcal{L}_i$ is the loss for the $i$-th sample.
+    \item $C$ is the number of possible classes.
+    \item $y_{i,j}$ represents the ground truth label for class $j$. It is a one-hot encoded vector, meaning:
+    \begin{itemize}
+        \item $y_{i,j} = 1$ if the sample belongs to class $j$.
+        \item $y_{i,j} = 0$ otherwise.
+    \end{itemize}
+    \item $\hat{y}_{i,j}$ is the predicted probability for class $j$, obtained from a softmax function.
+    \item $\log(\hat{y}_{i,j})$ takes the logarithm of the predicted probability.
+\end{itemize}
+
+We also looked at the learning curves of the deep leaning models (BERT, AttentionNN) to assess the overfit of each one and compared the models in terms of 
+accuracy for functions of different complexity obtaining interesting results revealing their strengths and weaknesses.
 
@@
 
-\begin{table}[h!]
+As we have mentioned, the models were mainly tested for their scalability with having more functions per author while decreasing the distinct author set size.
+The following table \ref{tab:comparison} summarizes how each model performed in terms of accuracy and the time required for training (until the validation loss stopped decreasing) across diverse datasets
+and figure \ref{fig:3d-bars} visualizes the achieved performances on Accuracy, Precision and Recall. From table \ref{tab:comparison} we can say that BERT 
+trained on tokens extracted from the raw source code performed the best at each of the author set sizes, especially with accuracy over 90\% on the smallest set, however with the longest required time to train the model, more than
+9.5 times longer than it took to train AttentionNN to its full potential and 36.4 times longer than the time required for Random Forests which scored just 4.8 percent less.
+What also interested us is that From \texttt{set\_size=11} to \texttt{set\_size=3}, BERT improved only by 0.86\% which is the smallest improvement across all the experiments and indicates that increasing the minimum function
+count per author from 50 to 58 had little impact on the model's ability to distinguish authors. 
+On the other hand, from being the worst, BERT fed with AST pre-order traversal scored a 40\% higher under these conditions.
+
+\begin{table}[H]
     \centering
@@
     Random Forests (TF--IDFs w/o comments) & 66.24\%\;|\;$3.63\,\mathrm{min}$ & 70.60\%\;|\;$0.23\,\mathrm{min}$ & 80.56\%\;|\;$0.08\,\mathrm{min}$ & 85.84\%\;|\;$0.05\,\mathrm{min}$ \\
-    BERT (source code w/o comments)        & 72.47\%\;|\;$70.33\,\mathrm{min}$ & 82.72\%\;|\;$32.49\,\mathrm{min}$ & 89.84\%\;|\;$9.34\,\mathrm{min}$  & 90.70\%\;|\;$1.82\,\mathrm{min}$ \\
-    BERT (AST pre\_order traversal)        & 24.07\%\;|\;$73.53\,\mathrm{min}$ & 35.66\%\;|\;$30.00\,\mathrm{min}$ & 40.62\%\;|\;$5.69\,\mathrm{min}$  & 83.72\%\;|\;$1.22\,\mathrm{min}$ \\
-    AttentionNN (AST paths)              & --- & --- & --- & --- \\
+    BERT (source code w/o comments)        & 72.47\%\;|\;$30.33\,\mathrm{min}$ & 82.72\%\;|\;$9.49\,\mathrm{min}$ & 89.84\%\;|\;$5.3\,\mathrm{min}$  & \textbf{90.70\%\;|\;\boldmath$1.82\,\mathrm{min}$}\\
+    BERT (AST pre\_order traversal)        & 24.07\%\;|\;$33.53\,\mathrm{min}$ & 35.66\%\;|\;$10.3\,\mathrm{min}$ & 40.62\%\;|\;$6.2\,\mathrm{min}$  & 83.72\%\;|\;$1.22\,\mathrm{min}$ \\
+    AttentionNN (AST NtN paths)            & 36.93\%\;|\;$3.53\,\mathrm{min}$ & 53.42\%\;|\;$1.19\,\mathrm{min}$ & 60.82\%\;|\;$0.54\,\mathrm{min}$  & 69.3\%\;|\;$0.19\,\mathrm{min}$ \\
     \hline
@@
 
-\begin{figure}[h!]
+\begin{figure}[H] 
+    \centering
+    \includegraphics[width=0.6\textwidth]{figures/conf_matrix_sourcode_bert_ts27.pdf}
+    \caption{Confusion matrix for best performative model BERT trained on author set of size 27.}
+    \label{fig:conf_matrix_bert}
+\end{figure}
+
+The confusion matrix shows, that the model performed best on 2 most represented authors
+(see \ref{fig:distr_names}). Although if we take 8 most represented authors, only 5 of them are also 
+the most accurately classified.
+
+In terms of overfit BERT with tokenized source codes was able to generalize the datasets in effective manner reaching 1.41, 1.24, 1.05, 1.12 for overfit ratios 
+calculated like this
+
+\begin{equation}
+    \text{Overfit Ratio} = \frac{\text{Train Accuracy}}{\text{Test Accuracy}}
+\end{equation}
+
+and only the latter two are under the limit of 1.2 - being optimally fitted.
+
+Random forests reached 1.51, 1.3, 1.24, 1.15 overfit ratios which still show a considerable amount of generalization. However AttentionNN was not that successful with generalizing 
+and with the dataset consisting of 3 distinct authors it got overfit of 1.4 and only increasing with each larger dataset, meaning insufficient data is being provided for the network so it starts memorizing particular
+patterns instead of actually generalizing above authors, which does not allow the model to perform better on unseen data affecting its validation accuracy.
+
+\begin{figure}[H]
 \centering
@@
 \addplot3+[ybar, bar width=12pt, fill=blue, draw=none, bar shift=-0.2]
-    coordinates {(1.1,4,85.84) (1,3,80.56) (1,2,70.60) (1,1,66.24)}; % Coordinates are adjusted to match the reversed ytick
+    coordinates {(1.1,4,85.84) (1,3,80.56) (1,2,70.60) (1,1.1,66.24)}; % Coordinates are adjusted to match the reversed ytick
 \addplot3+[ybar, bar width=12pt, fill=red, draw=none, bar shift=0]
-    coordinates {(1.1,4,89.36) (1,3,84.56) (1,2,79.27) (1,1,71.73)}; % Coordinates are adjusted to match the reversed ytick
+    coordinates {(1.1,4,89.36) (1,3,84.56) (1,2,79.27) (1,1.1,71.73)}; % Coordinates are adjusted to match the reversed ytick
 \addplot3+[ybar, bar width=12pt, fill=brown, draw=none, bar shift=+0.2]
-    coordinates {(1.1,4,84.96) (1,3,72.03) (1,2,69.73) (1,1,65.55)}; % Coordinates are adjusted to match the reversed ytick
+    coordinates {(1.1,4,84.96) (1,3,72.03) (1,2,69.73) (1,1.1,65.55)}; % Coordinates are adjusted to match the reversed ytick
 
@@
     mark=none,
-    blue,
+    purple,
     dotted,
-    line width=1pt,
+    line width=2pt,
     nodes near coords,
@@
         mark=none,
-        fill=none,
+        fill=gray!30,
         draw=none,
@@
         mark=none,
-        blue,
+        purple,
         dotted,
-        line width=1pt,
+        line width=2pt,
         nodes near coords,
@@
             mark=none,
-            fill=none,
+            fill=gray!30,
             draw=none,
@@
             font=\scriptsize,
-            yshift=0pt
+            yshift=0pt,
+            xshift=-10pt
         }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\foreach \yy in {1,2,3,4} {
-    \addplot3+[ybar, bar width=12pt, fill=blue, draw=none, bar shift=-0.2] coordinates {(4,\yy,0)};
-    \addplot3+[ybar, bar width=12pt, fill=red, draw=none, bar shift=0] coordinates {(4,\yy,0)};
-    \addplot3+[ybar, bar width=12pt, fill=brown, draw=none, bar shift=+0.2] coordinates {(4,\yy,0)};
+\addplot3+[ybar, bar width=12pt, fill=blue, draw=none, bar shift=-0.2]
+    coordinates {(4,3.9,69) (4,3,60.82) (4,2,53.42) (4,1,36.93)}; % Coordinates are adjusted to match the reversed ytick
+\addplot3+[ybar, bar width=12pt, fill=red, draw=none, bar shift=0]
+    coordinates {(4,3.9,71.25) (4,3,69.6) (4,2,62.6) (4,1,49.73)}; % Coordinates are adjusted to match the reversed ytick
+\addplot3+[ybar, bar width=12pt, fill=brown, draw=none, bar shift=+0.2]
+    coordinates {(4,3.9,68.76) (4,3,60.61) (4,2,53.45) (4,1,36.34)}; % Coordinates are adjusted to match the reversed ytick
+\foreach \yA/\zA/\yB/\zB in {
+    4/69.3/3/60.82,
+    3/60.82/2/53.42,
+    2/53.49/1/36.93
+} {
+    \addplot3+[
+        mark=none,
+        purple,
+        dotted,
+        line width=2pt,
+        nodes near coords,
+        every node near coord/.append style={
+            mark=none,
+            fill=gray!30,
+            draw=none,
+            text=black,
+            font=\scriptsize,
+            yshift=0pt,
+            xshift=-8pt
+        }
+    ] coordinates {
+        (4,\yA,\zA) (4,\yB,\zB)
+    };
 }
@@
 \caption{Bar chart comparing Accuracy (with accuracy values labeled), Precision, and Recall across models and author set sizes with minimum function count per author. }
-\label{fig:3d-bars-legend-brown}
+\label{fig:3d-bars}
 \end{figure}
 
+\subsection{Comparing models' performance depending on the complexity of function}
+
+In our task of identifying an author given a single function, a natural question arises: how did the models 
+perform on functions of different complexities? The basic indicator of function complexity is its length. So we came up with strategy, 
+how to visually compare models, even when the number of each model's misclassifications is different. We scaled the models histogram bin values by 
+an index that is obtained like this:
+
+\begin{equation}
+    \text{index}_{i} = \frac{\text{Number of misclassifications of the worst performative model}}{\text{Number of misclassifications of model}_{i}}
+\end{equation}
+
+This way we obtain visual representation that objectively accounts for the fact that the number of errors is different.
+In the following figure we visualized the frequencies of misclassified function lengths in tokens for Random Forests, BERT and AttentionNN scaling each 
+model with its index.
+
+\begin{figure}[H] 
+    \centering
+    \includegraphics[width=0.8\textwidth]{figures/misclass_lens.pdf}
+    \caption{The frequency distribution of misclassified function lengths in tokens for Random Forests, BERT and AttentionNN}
+    \label{fig:misclass_lens}
+\end{figure}
+
+From the figure \ref{fig:misclass_lens} the only statement we can say is that the model AttentionNN particularly struggled with functions of length lower than 40 tokens
+but did quite well on the longer ones. This will be discussed in the following subsection more deeply \ref{subsec:comp_over_quant}.
+Other models performance is not really distinguishable from this visualization, so we performed the same index strategy with the Halstead complexity metrics.
+Measuring each misclassified function's Halstead complexity and plotting it for Random Forests and BERT trained on tokenized source code, we obtained following figure.
+
+\begin{figure}[H] 
+    \centering
+    \includegraphics[width=0.8\textwidth]{figures/misclass_hs.png}
+    \caption{The frequency distribution of misclassified function Halstead complexity for Random Forests and BERT.}
+    \label{fig:misclass_halstead}
+\end{figure}
+
+This error distribution reveals that Random Forests struggled more with Halstead complexity increasing from the marked red point compared to the BERT 
+Specifically there are 1.59 times more misclassifications made by the Random Forest classifier than there are by BERT measured from the red point.
+However, closer we look to 0 on x axis, the more errors we see made by BERT. This would be the main difference between the deep learning and 
+machine learning method. The deep learning method utilizes self-attention that enables it to capture more complex relationships but struggles
+to categorize when limited or not complex enough data is provided, not being able to find the relationships that characterize the author of the function. 
+On the other hand, the machine learning TF-IDF method struggles with the more complex data because the representation lacks context understanding and solely 
+relies on the number of occurrences in the sample.
+
+\subsection{Attention neural network - complexity over quantity}  \label{subsec:comp_over_quant}
+
+Considering the observation in figure \ref{fig:misclass_lens}, where AttentionNN performed worst on lengths between 20 and 40 tokens,
+we assesed the model in a deeper manner. Evaluating accuracies across functions distributed into bins according to 2 metrics. 
+Number of nodes in the AST and depth of the AST of the function.
+
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}{0.48\textwidth} 
+        \centering
+        \includegraphics[width=\linewidth]{figures/attention_nn/ann_acc_num_nodes.pdf}
+        \caption{Accuracy across different bins of number of nodes in function's AST}
+        \label{fig:distr_names}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.48\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{figures/attention_nn/ann_acc_depth.pdf}
+        \caption{Accuracy across different bins of depth in function's AST}
+        \label{fig:disrt_lens}
+    \end{subfigure}
+    \caption{AttNN Accuracy of functions with different complexities}
+    \label{fig:eda_distr}
+\end{figure}
+
+From the figure we can deduce, there is an emerging trend in accuracy from bins of lower value to the bins of higher value.
+Eventhough the frequency distribution of both variables indicates more frequency on the lower values we observe a lower
+accuracy. This tells that the model is learning rather from more structural complex functions than simple non-nesting and 
+control-flow divergent ones.
+
 \section{Discussion}
-Here we discuss the implications of our results.
+
+Our findings show that deep learning models, especially BERT, perform better than traditional machine learning methods for authorship attribution 
+of single functions. But this comes at a cost—BERT needs way more computation and training time. On the other hand, Random Forest with TF-IDF 
+tokenization held up surprisingly well, especially with larger author sets, and trained very fast.
+
+A key takeaway is that BERT trained on raw source code tokens consistently had the best accuracy across all dataset sizes. However, 
+its improvement from \texttt{set\_size=11} to \texttt{set\_size=3} was only 0.86\%, meaning increasing the minimum function count per author 
+after a certain point doesn’t do much. Interestingly, BERT using AST pre-order traversal struggled with large author sets but got way better 
+when we reduced the dataset to three authors. This shows how important it is to choose the right input representation for deep learning models.
+
+Looking at misclassifications, each model had its weak points. BERT had trouble with short functions that had low Halstead complexity— they 
+don’t provide enough meaningful context to determine the author. Random Forests, which rely more on statistical patterns than contextual 
+understanding, started failing as function complexity increased. This makes sense since TF-IDF representations capture token frequency but 
+don’t really understand code structure.
+
+AttentionNN, which works with AST-derived node-to-node paths, preferred structurally complex functions over simple ones. This means that 
+networks using structural information may do better with more complex functions rather than short, flat ones. However, AttentionNN didn’t 
+generalize well when data was limited, and it started overfitting when trained on small author sets. This shows we either need bigger 
+datasets or more regularization to prevent overfitting.
+
+The results highlight a clear trade-off between model complexity, training time, and accuracy. Deep learning models, while powerful, 
+require serious computational resources, making them harder to scale in real-world use cases. Random Forests might not be as accurate, 
+but they’re super efficient and a solid alternative when speed is a concern.
+
+Going forward, it would be interesting to combine these methods. For example a hybrid approach using BERT embeddings with Random Forest 
+classifier could balance accuracy and efficiency. Also, using graph-based neural networks to process ASTs more effectively might push 
+authorship attribution even further. Exploring data augmentation techniques or better regularization strategies could also help deep 
+learning models generalize better, especially with AST-based inputs.
 
 \section{Conclusion}
-We conclude that this document compiles successfully. If you can read all sections, view the figure, and see the mathematical expressions, your LaTeX environment is correctly set up.
+
+In this study, we compared traditional machine learning models and deep learning models for authorship attribution of individual functions. 
+Our results showed that while BERT-based models provided the highest accuracy, they required significantly longer training times and 
+computational resources. In contrast, Random Forests trained on TF-IDF representations performed surprisingly well given their simplicity 
+and efficiency.
+
+The choice of model largely depends on the trade-off between accuracy and efficiency. If the goal is purely high accuracy and computational 
+resources aren’t a constraint, then BERT trained on tokenized source code is the best option. However, if speed and efficiency are more 
+important, Random Forests provide a strong alternative with competitive accuracy.
+
+We also found that different function representations significantly affect model performance. BERT struggled with AST-based pre-order 
+traversal when the number of authors was high.
+
+AttentionNN performed better on more structurally complex functions rather than shorter ones. This highlights the need for 
+careful feature selection capturing the complex relationships when applying deep learning to authorship attribution.
+
+Overall, this study demonstrates that while deep learning methods achieve state-of-the-art performance, traditional 
+machine learning approaches still hold value, especially in resource-constrained environments.
+\bibliographystyle{plain} 
+\bibliography{references}
 

### 2025-03-13 article.tex
ieee version started

@@
 The corresponding AST is structured as follows:
-\begin{verbatim}
-TranslationUnit
-|--FunctionDefinition `int add(int a, int b) {`
-   |--BasicTypeSpecifier `int`
-   |--FunctionDeclarator `add(`
-   |  |--IdentifierDeclarator `add`
-   |  |--ParameterSuffix `(int a, int b)`
-   |     |--ParameterDeclaration `int a`
-   |     |  |--BasicTypeSpecifier `int`
-   |     |  |--IdentifierDeclarator `a`
-   |     |--ParameterDeclaration `int b`
-   |        |--BasicTypeSpecifier `int`
-   |        |--IdentifierDeclarator `b`
-   |--CompoundStatement `{ return a + b; }`
-      |--ReturnStatement `return a + b;`
-         |--AddExpression `a + b`
-            |--IdentifierName `a`
-            |--IdentifierName `b`
-\end{verbatim}
-\label{verbatim:AST_example}
 

### 2025-03-13 myIEEE/article.tex
ieee version started

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts

% Required packages for IEEEtran
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}

% Additional packages from your article (ensuring they do not interfere with IEEE margins)
\usepackage{subcaption}
\usepackage{pgfplots}
\pgfplotsset{compat=1.17}
\usepackage{tikz}
\usepackage{listings}
\usepackage{verbatim}
\usepackage[edges]{forest}
\usepackage{adjustbox}

% Do not load geometry or bookmark in IEEEtran
%\usepackage{geometry}
%\usepackage{bookmark}

\begin{document}

\title{Comparing Machine Learning and Deep Learning Models with Attention Mechanisms for Multiple Authorship Attribution\\
\thanks{Identify applicable funding agency here. If none, delete this.}
}

\author{
\IEEEauthorblockN{Richard Čerňanský}
\IEEEauthorblockA{\textit{Department Name}\\
\textit{Organization}\\
City, Country\\
email@example.com}
\and
\IEEEauthorblockN{Ing. Juraj Petrík}
\IEEEauthorblockA{\textit{Department Name}\\
\textit{Organization}\\
City, Country\\
email@example.com}
}

\maketitle

\begin{abstract}
Authorship attribution (AA) in source code is crucial for plagiarism detection, forensic analysis, and software maintenance—especially when author information is missing or ambiguous. This field is not as well explored as written communication AA, and there are new models to be tested on different sized datasets. This study compares traditional machine learning models, such as Random Forests trained on TF-IDF tokenized representations, with deep learning models, including BERT and an Attention-based Neural Network (AttentionNN), for the task of multiple authorship attribution. We evaluate different representations of source code, including raw tokenized text and Abstract Syntax Tree (AST) structures, to determine their effectiveness in distinguishing authorship. The models are tested on a dataset of 3,004 C language functions from Google Code Jam spanning 2009--2018, assessing their accuracy, precision, recall, and training efficiency across varying author set sizes. Our findings highlight the strengths and limitations of each approach, showing that while deep learning models, particularly BERT trained on raw source code, achieve the highest accuracy, Random Forests offer competitive results with significantly lower computational cost. The study also explores the impact of function complexity on model performance, revealing that AttentionNN benefits from structural complexity while struggling with short functions. The results emphasize the trade-offs between accuracy, computational efficiency, and data representation, providing insights into the optimal model selection for source code authorship attribution.
\end{abstract}

\begin{IEEEkeywords}
Authorship attribution, machine learning, deep learning, BERT, attention mechanisms, source code.
\end{IEEEkeywords}

\section{Introduction}

Authorship attribution in source code analysis is a crucial task with applications in plagiarism detection, forensic investigations, 
and intellectual property protection, as well as software maintenance and quality analysis when author information is missing, inaccurate, 
or ambiguous. Given a function written in a programming language, determining its author requires capturing distinct stylistic patterns. 
Traditional machine learning (ML) methods, such as Random Forests trained on TF-IDF tokenized representations, have shown promise in text-based 
authorship identification. However, recent advancements in deep learning (DL) have introduced transformer-based models, such as BERT, which leverage 
contextual embeddings and self-attention mechanisms to extract more meaningful patterns from textual data.

Despite the success of deep learning models in natural language processing (NLP), their application to source code authorship 
attribution across different dataset sizes remains relatively unexplored. Unlike natural language, source code exhibits a more 
rigid syntax and structure, making its representation a key challenge. Several approaches have been proposed—from direct tokenized 
representations of raw source code to abstract syntax tree (AST) transformations that capture program logic—but it remains unclear which 
representation is most effective for distinguishing authorship across various levels of dataset complexity.

This study systematically compares machine learning and deep learning models, including Random Forests, BERT trained on tokenized source code, 
BERT trained on AST representations, and an Attention-based Neural Network (AttentionNN) utilizing AST-derived node-to-node paths. The evaluation 
is conducted on a dataset of 3,004 C language functions from Google Code Jam (GCJ) spanning 2009 to 2018 \cite{gcj}. The models are assessed 
based on their accuracy, precision, recall, and training efficiency across different author set sizes. Through this comparison, we aim to 
determine which model best balances accuracy and computational efficiency, as well as the effectiveness of different function representations 
for authorship attribution.

\section{Methods}

\subsection{Source Code Representation}

In \cite{information}, the authors present a division of features based on the extraction method:
\begin{itemize}
    \item \textbf{Stylometric features:} Capture structural characteristics that represent an author’s preference for various statements, keywords, 
    and nesting sizes.
    \item \textbf{N-gram features:} Extract contiguous sequences of bytes, characters, or words using a fixed-length sliding window.
    \item \textbf{Graph-based features:} Capture underlying traits in the code such as deep nesting, control flow, and data flow. This includes 
    representations such as the AST, Program Dependency Graph (PDG), Control Flow Graph (CFG), and RFG.
    \item \textbf{Behavioral features:} Focus on the program’s runtime behavior (e.g., CPU and memory usage) and instruction-level features from binaries.
    \item \textbf{Neural Network Generated Representations:} Recently, neural networks have been used to generate embeddings of source code. This approach 
    is applied in this study by using AST analysis to provide input for the neural network.
\end{itemize}

\subsection{Abstract Syntax Tree (AST)}

Based on the definition in \cite{sun2023AST}, an Abstract Syntax Tree (AST) is a tree-based representation of source code where nodes represent elements of the code and edges represent the relationships between these elements. The AST is an essential component in the compilation process, serving as an intermediate representation used for optimization and machine code generation. Nodes in the AST can be either internal (non-leaf), which define the program's constructs, or leaf nodes, which store actual textual values.

Below is an example of a simple C function and its corresponding AST:

\begin{figure}[!t]
    \centering
    % C code listing in a minipage
    \begin{minipage}{\columnwidth}
    \begin{lstlisting}[caption={Simple C function for addition}, label={lst:add}, basicstyle=\footnotesize\ttfamily, frame=single, numbers=left]
    int add(int a, int b) {
        return a + b;
    }
    \end{lstlisting}
    \end{minipage}

    \vspace{1em} % Adjust vertical space between listing and diagram

    % AST diagram scaled to column width
    \resizebox{\columnwidth}{!}{%
    \begin{forest}
    for tree={
      font=\fontsize{14pt}{16pt}\selectfont,
      l sep=3cm,
    }
    [TranslationUnit
      [\textbf{FunctionDefinition}: {int add(int a, int b) \{}
        [\textbf{BasicTypeSpecifier}: {int}]
        [\textbf{FunctionDeclarator}: {add(}
          [\textbf{IdentifierDeclarator}: {add}]
          [\textbf{ParameterSuffix}: {(int a, int b)}
            [\textbf{ParameterDeclaration}: {int a}
              [\textbf{BasicTypeSpecifier}: {int}]
              [\textbf{IdentifierDeclarator}: {a}]
            ]
            [\textbf{ParameterDeclaration}: {int b}
              [\textbf{BasicTypeSpecifier}: {int}]
              [\textbf{IdentifierDeclarator}: {b}]
            ]
          ]
        ]
        [\textbf{CompoundStatement}: {\{ return a + b; \}}
          [\textbf{ReturnStatement}: {return a + b;}
            [\textbf{AddExpression}: {a + b}
              [\textbf{IdentifierName}: {a}]
              [\textbf{IdentifierName}: {b}]
            ]
          ]
        ]
      ]
    ]
    \end{forest}
    }
    \caption{(a) C function code listing and (b) corresponding AST diagram.}
    \label{fig:combined_ast}
\end{figure}

\subsubsection{AST --- Types of Features}

Features extracted from the AST fall into four main categories:
\begin{itemize}
    \item \textbf{Structural:} Capture the AST's complexity (e.g., depth, number of nodes, average branching factor).
    \item \textbf{Semantic:} Reflect semantic information such as the distribution of node types or distinct root-to-leaf paths.
    \item \textbf{Syntactic:} Describe the programming paradigm and logical complexity (e.g., number of functions, control flow units).
    \item \textbf{Combined:} Use-case specific features that combine the aforementioned strategies for tasks like authorship attribution or 
    function name classification.
\end{itemize}

\subsubsection{Halstead Complexity Measures}

To evaluate function complexity, we used Halstead complexity measures (Difficulty, Volume, Effort) as defined in \cite{halstead}. 
These measures were aggregated into a single value using a weighted sum with ratios 4:3:3. (Halstead length was not incorporated, as it was analyzed separately.)

\subsection{Our Data Representation}

For training the models, a prediction task data pipeline was constructed as follows:

\begin{figure}[!t]
    \centering
    \includegraphics[width=2.5cm,height=3.5cm]{figures/high_level_prediciton_task.png}
    \caption{\cite{compiler_based} High-level overview of the function classification prediction task.}
    \label{fig:predictive_task}
\end{figure}

Various representations of each function’s raw source code were utilized:
\begin{itemize}
    \item Word-tokenized TF-IDF vectors for Random Forests.
    \item Byte Pair Encoding (BPE) of raw source code for BERT.
    \item Linearized AST pre-order traversal for BERT.
    \item A set of 600 randomly selected node-to-node paths for AttentionNN.
\end{itemize}
The PsycheC project's C language frontend parser \cite{psychec} was used to convert code into ASTs. Models were trained in the 
TensorFlow.Keras framework using different author set sizes (110, 27, 11, and 3) with minimum function counts per author of 25, 40, 50, and 58 respectively.

\subsection{Data Preprocessing \& Exploratory Data Analysis}

The dataset consisted of thousands of solutions from hundreds of authors. From each solution, the function and its author were extracted. 
The largest dataset comprised $3\,004$ C language functions from GCJ solutions (2009--2018). Preprocessing involved duplicate removal,
elimination of authors with disproportionately many samples, and exclusion of functions that could not be parsed into an AST. Frequency 
distributions of authors and function lengths were analyzed for further insights.

\begin{table*}[!t]
    \renewcommand{\arraystretch}{1.3}
    \caption{Comparison of models across author attribution accuracy and training time for different author set sizes.}
    \label{tab:comparison}
    \centering
    \begin{tabular}{lcccc}
    \hline
    \multicolumn{5}{c}{\textbf{Model Performance on Author Set Sizes}} \\
    \hline
    \textbf{Model} & \textbf{110} & \textbf{27} & \textbf{11} & \textbf{3} \\
    \hline
    Random Forests (TF-IDF) & 66.24\% / 3.63 min & 70.60\% / 0.23 min & 80.56\% / 0.08 min & 85.84\% / 0.05 min \\
    BERT (source code)    & 72.47\% / 30.33 min & 82.72\% / 9.49 min & 89.84\% / 5.30 min  & \textbf{90.70\% / 1.82 min} \\
    BERT (AST traversal)  & 24.07\% / 33.53 min & 35.66\% / 10.30 min & 40.62\% / 6.20 min  & 83.72\% / 1.22 min \\
    AttentionNN (AST paths)& 36.93\% / 3.53 min  & 53.42\% / 1.19 min  & 60.82\% / 0.54 min  & 69.30\% / 0.19 min \\
    \hline
    \end{tabular}
\end{table*}

\begin{figure}[!t]
    \centering
    \begin{subfigure}{0.48\textwidth}
        \centering
        \includegraphics[width=\linewidth]{figures/distribution_names.pdf}
        \caption{Frequency distribution of author names}
        \label{fig:distr_names}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\textwidth}
        \centering
        \includegraphics[width=\linewidth]{figures/distr_lens_tokens.pdf}
        \caption{Frequency distribution of function lengths}
        \label{fig:disrt_lens}
    \end{subfigure}
    \caption{Important frequency distributions of our dataset.}
    \label{fig:eda_distr}
\end{figure}

Figure~\ref{fig:distr_names} shows that most function lengths are between 10 and 50 tokens, 
making single-function authorship attribution challenging when only short code snippets are available.

\section{Models Setup \& Training}

Random Forests and AttentionNN models were trained using a stratified cross-validation strategy (5 folds: 3 for training,
 1 for validation, and 1 for testing). Because BERT training requires considerably more time, it was repeated three times, with evaluation metric variance below 1\%.

\subsection{Random Forests}
Random Forests were chosen as a baseline for text-based authorship attribution. They were trained on word-level tokens (using the regex \verb|r"\b\w+\b"|). A grid search was performed to optimize hyperparameters:
\begin{itemize}
    \item \textbf{n\_estimators}: \{50, 100, 200, 500\}
    \item \textbf{max\_depth}: \{10, 20, 30, None\}
    \item \textbf{min\_samples\_split}: \{2, 5, 10\}
    \item \textbf{min\_samples\_leaf}: \{1, 2, 4\}
    \item \textbf{max\_features}: \{"sqrt", "log2", None\}
    \item \textbf{bootstrap}: \{True, False\}
\end{itemize}
The best hyperparameters were then used to train the final model.

\subsection{BERT}
For both BERT setups, Microsoft’s pretrained \texttt{graphcodebert-base} \cite{codebert} was used. The graphcodebert tokenizer uses Byte Pair Encoding, which is suitable for programming languages with rigid syntax. Comments were removed to ensure fairness when comparing models that utilize ASTs. BERT was trained on two distinct representations: tokenized raw source code and tokenized AST pre-order traversal.

\subsection{Attention Neural Network}
The AttentionNN model leverages AST-derived node-to-node paths (path-contexts) as proposed in \cite{code2vec}. Each path is represented as a triplet (start node, path of node kinds, terminal node) and encoded into vector space. The model then learns to weigh these contexts via an attention mechanism to distinguish authors.

\subsection{Evaluation Methods}
Evaluation metrics included validation accuracy, precision, recall, training time, and Halstead complexity measures. Deep learning models were trained with the categorical cross-entropy loss function:
\begin{equation}
    \mathcal{L}_i = - \sum_{j=1}^{C} y_{i,j} \log(\hat{y}_{i,j})
\end{equation}
where:
\begin{itemize}
    \item $\mathcal{L}_i$ is the loss for sample $i$.
    \item $C$ is the number of classes.
    \item $y_{i,j}$ is the one-hot encoded ground truth for class $j$.
    \item $\hat{y}_{i,j}$ is the predicted probability for class $j$ (obtained from a softmax function).
\end{itemize}
Learning curves were analyzed to assess overfitting. The overfit ratio was calculated as:
\begin{equation}
    \text{Overfit Ratio} = \frac{\text{Train Accuracy}}{\text{Test Accuracy}}
\end{equation}

\begin{figure*}[!t]
    \centering
    \begin{tikzpicture}
    \begin{axis}[
        view={60}{30},
        z buffer=sort,
        xlabel={Model},
        xlabel style={xshift=-40pt,yshift=40pt,align=center},
        ylabel={Author set size},
        ylabel style={xshift=25pt,yshift=20pt,align=center},
        zlabel={Metric (\%)},
        width=14cm,
        height=8cm,
        xtick={1,2,3,4},
        xticklabels={RF (TF-IDF), BERT (source code), BERT (AST pre\_order), AttentionNN},
        xticklabel style={font=\footnotesize,xshift=-20pt,yshift=10pt,align=center},
        ytick={4,3,2,1},
        yticklabels={3 (min. count 58), 11 (min. count 50), 27 (min. count 40), 110 (min. count 25)},
        yticklabel style={font=\footnotesize,xshift=25pt,yshift=5pt,align=center},
        zmin=0, zmax=100,
        grid=major,
        legend style={at={(0.5,1.15)},anchor=north,legend columns=3}
    ]

    % MODEL 1: RF (x = 1)
    \addplot3+[ybar, bar width=12pt, fill=blue, draw=none, bar shift=-0.2]
        coordinates {(1.1,4,85.84) (1,3,80.56) (1,2,70.60) (1,1.1,66.24)};
    \addplot3+[ybar, bar width=12pt, fill=red, draw=none, bar shift=0]
        coordinates {(1.1,4,89.36) (1,3,84.56) (1,2,79.27) (1,1.1,71.73)};
    \addplot3+[ybar, bar width=12pt, fill=brown, draw=none, bar shift=+0.2]
        coordinates {(1.1,4,84.96) (1,3,72.03) (1,2,69.73) (1,1.1,65.55)};
    \addplot3+[mark=none, purple, dotted, line width=2pt, nodes near coords,
        every node near coord/.append style={mark=none, fill=gray!30, draw=none, text=black, font=\scriptsize,yshift=0pt}]
        coordinates {(0.8,4,85.84) (0.8,3,80.56) (0.8,2,70.60) (0.8,1,66.24)};

    % MODEL 2: BERT (source code) (x = 2)
    \addplot3+[ybar, bar width=12pt, fill=blue, draw=none, bar shift=-0.2]
        coordinates {(2,4,90.70) (2,3,89.84) (2,2,82.72) (2,1,72.47)};
    \addplot3+[ybar, bar width=12pt, fill=red, draw=none, bar shift=0]
        coordinates {(2,4,93.33) (2,3,90.33) (2,2,84.66) (2,1,76.69)};
    \addplot3+[ybar, bar width=12pt, fill=brown, draw=none, bar shift=+0.2]
        coordinates {(2,4,91.67) (2,3,89.92) (2,2,82.25) (2,1,71.76)};
    \addplot3+[mark=none, purple, dotted, line width=2pt, nodes near coords,
        every node near coord/.append style={mark=none, fill=gray!30, draw=none, text=black, font=\scriptsize,yshift=0pt}]
        coordinates {(1.8,4,90.70) (1.8,3,89.84) (1.8,2,82.72) (1.8,1,72.47)};

    % MODEL 3: BERT (AST pre_order) (x = 3)
    \addplot3+[ybar, bar width=12pt, fill=blue, draw=none, bar shift=-0.2]
        coordinates {(3,4,83.72) (3,3,40.62) (3,2,35.66) (3,1,24.07)};
    \addplot3+[ybar, bar width=12pt, fill=red, draw=none, bar shift=0]
        coordinates {(3,4,84.72) (3,3,40.53) (3,2,36.41) (3,1,27.40)};
    \addplot3+[ybar, bar width=12pt, fill=brown, draw=none, bar shift=+0.2]
        coordinates {(3,4,82.58) (3,3,38.30) (3,2,33.93) (3,1,23.33)};
    \foreach \yA/\zA/\yB/\zB in {
        4/83.72/3/40.62,
        3/40.62/2/35.66,
        2/35.66/1/24.07
    } {
        \addplot3+[mark=none, purple, dotted, line width=2pt, nodes near coords,
            every node near coord/.append style={mark=none, fill=gray!30, draw=none, text=black, font=\scriptsize,yshift=0pt,xshift=-10pt}]
        coordinates {(3,\yA,\zA) (3,\yB,\zB)};
    }

    % MODEL 4: AttentionNN (x = 4)
    \addplot3+[ybar, bar width=12pt, fill=blue, draw=none, bar shift=-0.2]
        coordinates {(4,3.9,69) (4,3,60.82) (4,2,53.42) (4,1,36.93)};
    \addplot3+[ybar, bar width=12pt, fill=red, draw=none, bar shift=0]
        coordinates {(4,3.9,71.25) (4,3,69.6) (4,2,62.6) (4,1,49.73)};
    \addplot3+[ybar, bar width=12pt, fill=brown, draw=none, bar shift=+0.2]
        coordinates {(4,3.9,68.76) (4,3,60.61) (4,2,53.45) (4,1,36.34)};
    \foreach \yA/\zA/\yB/\zB in {
        4/69.3/3/60.82,
        3/60.82/2/53.42,
        2/53.49/1/36.93
    } {
        \addplot3+[mark=none, purple, dotted, line width=2pt, nodes near coords,
            every node near coord/.append style={mark=none, fill=gray!30, draw=none, text=black, font=\scriptsize,yshift=0pt,xshift=-8pt}]
        coordinates {(4,\yA,\zA) (4,\yB,\zB)};
    }

    % LEGEND
    \addplot3+[ybar, fill=blue, draw=none, bar width=12pt, legend image code/.code={
        \draw[fill=blue] (0cm,-0.1cm) rectangle (0.6cm,0.1cm);
    }] coordinates {(0,0,-9999)};
    \addlegendentry{Accuracy}

    \addplot3+[ybar, fill=red, draw=none, bar width=12pt, legend image code/.code={
        \draw[fill=red] (0cm,-0.1cm) rectangle (0.6cm,0.1cm);
    }] coordinates {(0,0,-9999)};
    \addlegendentry{Precision}

    \addplot3+[ybar, fill=brown, draw=none, bar width=12pt, legend image code/.code={
        \draw[brown, mark=x] plot coordinates {(0cm,0cm)};
    }] coordinates {(0,0,-9999)};
    \addlegendentry{Recall}

    \end{axis}
    \end{tikzpicture}
    \caption{3D bar chart comparing Accuracy, Precision, and Recall across models and author set sizes.}
    \label{fig:3d-bars}
\end{figure*}

\section{Results}

The models were evaluated for scalability as the number of functions per author increased while the distinct author set size decreased. Table~\ref{tab:comparison} summarizes model performance (accuracy and training time) for various author set sizes.

Figure~\ref{fig:conf_matrix_bert} shows the confusion matrix for the best performing model (BERT on raw source code) for an author set size of 27.

\begin{figure}[!b]
    \centering
    \hspace*{-25pt}\includegraphics[width=1.2\columnwidth]{figures/misclass_lens.pdf}
    \caption{Frequency distribution of misclassified function lengths (in tokens) for Random Forests, BERT, and AttentionNN.}
    \label{fig:misclass_lens}
\end{figure}

We also analyzed the effect of function complexity. Figure~\ref{fig:misclass_lens} visualizes the frequency distribution of misclassified function lengths (in tokens) for Random Forests, BERT, and AttentionNN after scaling each model’s misclassification count by an index:
\begin{equation}
    \text{index}_{i} = \frac{\text{Number of misclassifications of the worst performing model}}{\text{Number of misclassifications of model } i}
\end{equation}

Figure~\ref{fig:misclass_halstead} shows the distribution of misclassified function Halstead complexity for Random Forests and BERT, revealing that while Random Forests struggle with higher complexity, BERT exhibits more misclassifications when complexity is low.

\begin{figure}[!b]
    \centering
    \hspace*{-25pt}\includegraphics[width=1.2\columnwidth]{figures/misclass_hs.png}
    \caption{Frequency distribution of misclassified function Halstead complexity for Random Forests and BERT.}
    \label{fig:misclass_halstead}
\end{figure}

Furthermore, we assessed the AttentionNN model’s performance based on function complexity by evaluating accuracy across bins of AST node count and AST depth. Figure~\ref{fig:eda_distr} presents subfigures for these two metrics, indicating that the model learns better from structurally complex functions.

A 3D bar chart (Figure~\ref{fig:3d-bars}) further visualizes Accuracy, Precision, and Recall across models and author set sizes. (The code for this chart uses TikZ and pgfplots.)

\section{Discussion}

Our findings indicate that deep learning models—especially BERT trained on raw source code—consistently outperform traditional methods like Random Forests in terms of accuracy, albeit at the expense of longer training times. Random Forests, despite their simplicity and efficiency, lag behind in accuracy when compared to BERT. Meanwhile, AttentionNN benefits from capturing structural complexity but tends to overfit with limited data. These results highlight the trade-offs between model complexity, training time, and overall performance. Future work might explore hybrid approaches (e.g., combining BERT embeddings with Random Forest classifiers) and improved regularization to balance accuracy and computational efficiency.

\section{Conclusion}

This study compared traditional machine learning models and deep learning approaches for authorship attribution of individual functions. While BERT-based models achieved state-of-the-art accuracy, they required significantly longer training times and computational resources compared to Random Forests. The choice of input representation (raw source code tokens versus AST-based traversal) greatly influenced performance. In resource-constrained environments, simpler models like Random Forests remain a viable option, whereas applications demanding higher accuracy may favor deep learning models despite their overhead. Future research could investigate hybrid models and advanced regularization techniques to further improve generalization.

\begin{figure}[!b]
    \centering
    % Scale to the column width (100%)
    \includegraphics[width=\columnwidth]{figures/conf_matrix_sourcode_bert_ts27.pdf}
    \caption{Confusion matrix for BERT trained on raw source code (author set size 27).}
    \label{fig:conf_matrix_bert}
\end{figure}

\begin{figure}[!b]
    \centering
    \begin{subfigure}{0.48\textwidth} 
        \centering
        \includegraphics[width=\linewidth]{figures/attention_nn/ann_acc_num_nodes.pdf}
        \caption{Accuracy vs. number of AST nodes}
        \label{fig:ann_acc_nodes}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\textwidth}
        \centering
        \includegraphics[width=\linewidth]{figures/attention_nn/ann_acc_depth.pdf}
        \caption{Accuracy vs. AST depth}
        \label{fig:ann_acc_depth}
    \end{subfigure}
    \caption{AttentionNN accuracy for functions with different structural complexities.}
    \label{fig:ann_acc}
\end{figure}

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}

### 2025-03-13 myIEEE/article.tex
jopp

@@
     \centering
-    \includegraphics[width=2.5cm,height=3.5cm]{figures/high_level_prediciton_task.png}
+    \includegraphics[width=3.5cm,height=5cm]{figures/high_level_prediciton_task.png}
     \caption{\cite{compiler_based} High-level overview of the function classification prediction task.}
@@
 
-\section{Discussion}
-
-Our findings indicate that deep learning models—especially BERT trained on raw source code—consistently outperform traditional methods like Random Forests in terms of accuracy, albeit at the expense of longer training times.
-Random Forests, despite their simplicity and efficiency, lag behind in accuracy when compared to BERT.
-Meanwhile, AttentionNN benefits from capturing structural complexity but tends to overfit with limited data.
-These results highlight the trade-offs between model complexity, training time, and overall performance.
-Future work might explore hybrid approaches (e.g., combining BERT embeddings with Random Forest classifiers) and improved regularization to balance accuracy and computational efficiency.
-
-\section{Conclusion}
-
-This study compared traditional machine learning models and deep learning approaches for authorship attribution of individual functions.
-While BERT-based models achieved state-of-the-art accuracy, they required significantly longer training times and computational resources compared to Random Forests.
-The choice of input representation (raw source code tokens versus AST-based traversal) greatly influenced performance.
-In resource-constrained environments, simpler models like Random Forests remain a viable option, whereas applications demanding higher accuracy may favor deep learning models despite their overhead.
-Future research could investigate hybrid models and advanced regularization techniques to further improve generalization.
-
-\begin{figure}[!b]
+\begin{figure}[!t]
     \centering
@@
 
+\section{Discussion}
+
+Our findings show that deep learning models, especially BERT, perform better than traditional machine learning methods for authorship attribution 
+of single functions. But this comes at a cost—BERT needs way more computation and training time. On the other hand, Random Forest with TF-IDF 
+tokenization held up surprisingly well, especially with larger author sets, and trained very fast.
+
+A key takeaway is that BERT trained on raw source code tokens consistently had the best accuracy across all dataset sizes. However, 
+its improvement from \texttt{set\_size=11} to \texttt{set\_size=3} was only 0.86\%, meaning increasing the minimum function count per author 
+after a certain point doesn’t do much. Interestingly, BERT using AST pre-order traversal struggled with large author sets but got way better 
+when we reduced the dataset to three authors. This shows how important it is to choose the right input representation for deep learning models.
+
+Looking at misclassifications, each model had its weak points. BERT had trouble with short functions that had low Halstead complexity— they 
+don’t provide enough meaningful context to determine the author. Random Forests, which rely more on statistical patterns than contextual 
+understanding, started failing as function complexity increased. This makes sense since TF-IDF representations capture token frequency but 
+don’t really understand code structure.
+
+AttentionNN, which works with AST-derived node-to-node paths, preferred structurally complex functions over simple ones. This means that 
+networks using structural information may do better with more complex functions rather than short, flat ones. However, AttentionNN didn’t 
+generalize well when data was limited, and it started overfitting when trained on small author sets. This shows we either need bigger 
+datasets or more regularization to prevent overfitting.
+
+The results highlight a clear trade-off between model complexity, training time, and accuracy. Deep learning models, while powerful, 
+require serious computational resources, making them harder to scale in real-world use cases. Random Forests might not be as accurate, 
+but they’re super efficient and a solid alternative when speed is a concern.
+
+Going forward, it would be interesting to combine these methods. For example a hybrid approach using BERT embeddings with Random Forest 
+classifier could balance accuracy and efficiency. Also, using graph-based neural networks to process ASTs more effectively might push 
+authorship attribution even further. Exploring data augmentation techniques or better regularization strategies could also help deep 
+learning models generalize better, especially with AST-based inputs.
+
 \begin{figure}[!b]
     \centering
-    \begin{subfigure}{0.48\textwidth} 
+    \begin{subfigure}{\columnwidth}
         \centering
@@
     \hfill
-    \begin{subfigure}{0.48\textwidth}
+    \begin{subfigure}{\columnwidth}
         \centering
@@
 
-\bibliographystyle{IEEEtran}
-\bibliography{references}
+\section{Conclusion}
+
+In this study, we compared traditional machine learning models and deep learning models for authorship attribution of individual functions. 
+Our results showed that while BERT-based models provided the highest accuracy, they required significantly longer training times and 
+computational resources. In contrast, Random Forests trained on TF-IDF representations performed surprisingly well given their simplicity 
+and efficiency.
+
+The choice of model largely depends on the trade-off between accuracy and efficiency. If the goal is purely high accuracy and computational 
+resources aren’t a constraint, then BERT trained on tokenized source code is the best option. However, if speed and efficiency are more 
+important, Random Forests provide a strong alternative with competitive accuracy.
+
+We also found that different function representations significantly affect model performance. BERT struggled with AST-based pre-order 
+traversal when the number of authors was high.
+
+AttentionNN performed better on more structurally complex functions rather than shorter ones. This highlights the need for 
+careful feature selection capturing the complex relationships when applying deep learning to authorship attribution.
+
+Overall, this study demonstrates that while deep learning methods achieve state-of-the-art performance, traditional 
+machine learning approaches still hold value, especially in resource-constrained environments.
+
+\begin{thebibliography}{00}
+    \bibitem{b1} G.
+Eason, B.
+Noble, and I.
+N.
+Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
+Trans.
+Roy.
+Soc.
+London, vol.
+A247, pp.
+529--551, April 1955.
+    \bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
+    \bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
+    \bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
+    \bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
+    \bibitem{b6} Y.
+Yorozu, M.
+Hirano, K.
+Oka, and Y.
+Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
+J.
+Magn.
+Japan, vol.
+2, pp.
+740--741, August 1987 [Digests 9th Annual Conf.
+Magnetics Japan, p.
+301, 1982].
+    \bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
+    \bibitem{b8} D. P. Kingma and M. Welling, ``Auto-encoding variational Bayes,'' 2013, arXiv:1312.6114. [Online]. Available: https://arxiv.org/abs/1312.6114
+    \bibitem{b9} S. Liu, ``Wi-Fi Energy Detection Testbed (12MTC),'' 2023, gitHub repository. [Online]. Available: https://github.com/liustone99/Wi-Fi-Energy-Detection-Testbed-12MTC
+    \bibitem{b10} ``Treatment episode data set: discharges (TEDS-D): concatenated, 2006 to 2009.'' U.S.
+Department of Health and Human Services, Substance Abuse and Mental Health Services Administration, Office of Applied Studies, August, 2013, DOI:10.3886/ICPSR30122.v2
+    \bibitem{b11} K. Eves and J. Valasek, ``Adaptive control for singularly perturbed systems examples,'' Code Ocean, Aug. 2023. [Online]. Available: https://codeocean.com/capsule/4989235/tree
+    \end{thebibliography}
 

### 2025-03-14 myIEEE/article.tex
referencese in progress

@@
 \usepackage{adjustbox}
+\usepackage{url}
 
@@
 
-\title{Comparing Machine Learning and Deep Learning Models with Attention Mechanisms for Multiple Authorship Attribution\\
+\title{Comparing Machine Learning and Deep Learning Models with Attention Mechanisms for Multiple Source Code Authorship Attribution\\
 \thanks{Identify applicable funding agency here. If none, delete this.}
@@
 \begin{abstract}
-Authorship attribution (AA) in source code is crucial for plagiarism detection, forensic analysis, and software maintenance—especially when author information is missing or ambiguous.
-This field is not as well explored as written communication AA, and there are new models to be tested on different sized datasets.
-This study compares traditional machine learning models, such as Random Forests trained on TF-IDF tokenized representations, with deep learning models, including BERT and an Attention-based Neural Network (AttentionNN), for the task of multiple authorship attribution.
-We evaluate different representations of source code, including raw tokenized text and Abstract Syntax Tree (AST) structures, to determine their effectiveness in distinguishing authorship.
-The models are tested on a dataset of 3,004 C language functions from Google Code Jam spanning 2009--2018, assessing their accuracy, precision, recall, and training efficiency across varying author set sizes.
-Our findings highlight the strengths and limitations of each approach, showing that while deep learning models, particularly BERT trained on raw source code, achieve the highest accuracy, Random Forests offer competitive results with significantly lower computational cost.
-The study also explores the impact of function complexity on model performance, revealing that AttentionNN benefits from structural complexity while struggling with short functions.
-The results emphasize the trade-offs between accuracy, computational efficiency, and data representation, providing insights into the optimal model selection for source code authorship attribution.
+    Authorship attribution (AA) in source code is crucial for plagiarism detection, 
+    forensic analysis, and software maintenance, especially when author 
+    information is missing or ambiguous. This field is not as well explored 
+    as the written communication AA and there are new models to be tested 
+    on different sized dataset. This study compares traditional 
+    machine learning models, such as Random Forests trained on TF-IDF 
+    tokenized representations, with deep learning models, including BERT 
+    and an Attention-based Neural Network (AttentionNN), for the task of 
+    multiple authorship attribution. We evaluate different representations 
+    of source code, including raw tokenized text and Abstract Syntax Tree (AST) 
+    structures, to determine their effectiveness in distinguishing authorship. 
+    The models are tested on various sized but relatively small datasets compared to the existing experiments, 
+    the largest consisting of 3,756 C language functions from 
+    Google Code Jam spanning 2009–2018, assessing their accuracy, precision, 
+    recall, and training efficiency across varying author set sizes. 
+    Our findings highlight the strengths and limitations of each approach, 
+    showing that while deep learning models, particularly BERT trained on 
+    raw source code, achieve the highest accuracy, Random Forests offer 
+    competitive results with significantly lower computational cost. 
+    The study also explores the impact of function complexity on model 
+    performance, revealing that AttentionNN benefits from structural 
+    complexity while struggling with short functions. The results emphasize 
+    the trade-offs between accuracy, computational efficiency, and data 
+    representation, providing insights into the optimal model selection 
+    for source code authorship attribution. 
 \end{abstract}
@@
 \begin{IEEEkeywords}
-Authorship attribution, machine learning, deep learning, BERT, attention mechanisms, source code.
+Authorship attribution, machine learning, deep learning, BERT, attention mechanisms, source code analysis.
 \end{IEEEkeywords}
@@
 
-Authorship attribution in source code analysis is a crucial task with applications in plagiarism detection, forensic investigations, 
-and intellectual property protection, as well as software maintenance and quality analysis when author information is missing, inaccurate, 
-or ambiguous. Given a function written in a programming language, determining its author requires capturing distinct stylistic patterns. 
-Traditional machine learning (ML) methods, such as Random Forests trained on TF-IDF tokenized representations, have shown promise in text-based 
-authorship identification. However, recent advancements in deep learning (DL) have introduced transformer-based models, such as BERT, which leverage 
-contextual embeddings and self-attention mechanisms to extract more meaningful patterns from textual data.
+Authorship attribution in source code analysis is a crucial task 
+with applications in plagiarism detection, forensic investigations, and 
+intellectual property protection, but also software maintenance 
+and software quality analysis \cite{pbnn}, where the information about an author 
+is missing, inaccurate or ambiguous.
+Given a function written in a programming
+language, determining its author requires capturing distinct stylistic 
+patterns. Traditional machine learning (ML) methods, such as Random 
+Forests trained on TF-IDF tokenized representations, have shown promise 
+in text-based authorship identification. However, recent advancements 
+in deep learning (DL) have introduced transformer-based models, such as 
+BERT, which leverage contextual embeddings and self-attention mechanisms 
+to extract more meaningful patterns from textual data.
 
-Despite the success of deep learning models in natural language processing (NLP), their application to source code authorship 
-attribution across different dataset sizes remains relatively unexplored. Unlike natural language, source code exhibits a more 
-rigid syntax and structure, making its representation a key challenge. Several approaches have been proposed—from direct tokenized 
-representations of raw source code to abstract syntax tree (AST) transformations that capture program logic—but it remains unclear which 
-representation is most effective for distinguishing authorship across various levels of dataset complexity.
+Despite the success of deep learning models in natural language 
+processing (NLP), their application to source code authorship attribution on different dataset sizes. 
+remains relatively unexplored. Unlike natural language, source code 
+exhibits a more rigid syntax and structure, making its representation 
+a key challenge. Several approaches have been proposed, ranging from 
+direct tokenized representations of the raw source code to abstract 
+syntax tree (AST) transformations that capture program logic. However, 
+it remains unclear which type of representation is most effective for 
+distinguishing authorship across different levels of dataset complexity.
 
-This study systematically compares machine learning and deep learning models, including Random Forests, BERT trained on tokenized source code, 
-BERT trained on AST representations, and an Attention-based Neural Network (AttentionNN) utilizing AST-derived node-to-node paths. The evaluation 
-is conducted on a dataset of 3,004 C language functions from Google Code Jam (GCJ) spanning 2009 to 2018 \cite{gcj}. The models are assessed 
-based on their accuracy, precision, recall, and training efficiency across different author set sizes. Through this comparison, we aim to 
-determine which model best balances accuracy and computational efficiency, as well as the effectiveness of different function representations 
-for authorship attribution.
+This study systematically compares machine learning and deep learning 
+models, including Random Forests, BERT trained on tokenized source code, 
+BERT trained on AST representations, and an Attention-based Neural Network 
+(AttentionNN) utilizing AST-derived node-to-node paths. The evaluation is 
+conducted on a dataset of 3,004 C language functions from Google Code Jam 
+(GCJ) spanning 2009 to 2018 \cite{gcj}. Most of the works in the field
+predict the authors based on files consisting of multiple language constructs, our 
+research however focuses on prediction of author given single function.
+The models are assessed based on their accuracy, precision, recall, and training efficiency across different 
+author set sizes.
+
+Through this comparison, we aim to determine which model best balances 
+accuracy and computational efficiency, as well as the effectiveness of 
+different function representations for authorship attribution.
 
@@
 
-In \cite{information}, the authors present a division of features based on the extraction method:
+In the paper \cite{information}, authors present division of features based on the method they use for extraction:
+
 \begin{itemize}
-    \item \textbf{Stylometric features:} Capture structural characteristics that represent an author’s preference for various statements, keywords, 
-    and nesting sizes.
-    \item \textbf{N-gram features:} Extract contiguous sequences of bytes, characters, or words using a fixed-length sliding window.
-    \item \textbf{Graph-based features:} Capture underlying traits in the code such as deep nesting, control flow, and data flow. This includes 
-    representations such as the AST, Program Dependency Graph (PDG), Control Flow Graph (CFG), and RFG.
-    \item \textbf{Behavioral features:} Focus on the program’s runtime behavior (e.g., CPU and memory usage) and instruction-level features from binaries.
-    \item \textbf{Neural Network Generated Representations:} Recently, neural networks have been used to generate embeddings of source code. This approach 
-    is applied in this study by using AST analysis to provide input for the neural network.
+
+    \item \textbf{Stylometric features:} 
+    They express the structural characters that represent the author’s preference for different statements, keywords, nesting sizes, etc.
+    \item \textbf{N-gram features:}
+    N-gram is a contiguous sequence of an n byte, character, or words extracted from the source. A sliding that has a fixed length generates 
+    the sequence.
+    \item \textbf{Graph based features:}
+    These capture underlying traits in the code such as deeply nesting code, control flow and data flow.
+    It includes : AST, Program Dependency Graph (PDG), Control Flow Graph (CFG), RFG.
+    \item \textbf{Behavioral features:}
+    This type completely different from the previous as it is not looking at the source code itself
+    but rather on the way that the program of the source code communicates with the computer.
+    In this category there are dynamic features which collect runtime measurements as CPU and memory usage.
+    Instruction features that are retrieved from binaries and reflect code properties like registers or immediate operands.
+    \item \textbf{Neural Networks Generated Representations of Source Code or Features:}
+    Lastly, in recent years, neural networks have been increasingly utilized 
+    to generate representations (such as embeddings) of source code. This 
+    approach is also applied in this thesis, using AST analysis to provide 
+    input for the neural network.
+
 \end{itemize}
@@
 
-Based on the definition in \cite{sun2023AST}, an Abstract Syntax Tree (AST) is a tree-based representation of source code where nodes represent elements of the code and edges represent the relationships between these elements.
-The AST is an essential component in the compilation process, serving as an intermediate representation used for optimization and machine code generation.
-Nodes in the AST can be either internal (non-leaf), which define the program's constructs, or leaf nodes, which store actual textual values.
+\quad Based on the full definition \cite{sun_ast} Abstract syntax tree (AST) is a tree-based representation of source code, where nodes represent various elements of the source code 
+and paths represent relationships of the structure. AST is an inseparable part of compilation process for many reasons. It is used as an 
+intermediate representation of program utilized for optimization and generation of machine code. Nodes of the tree can be either internal 
+(non-leaf) or leafs. Internal nodes define the program's constructs or operations for their children. Leaf nodes store the actual textual value.
 
-Below is an example of a simple C function and its corresponding AST:
+In the figure \ref{fig:ast_struc_example} is an example of a simple C function and its corresponding Abstract Syntax Tree (AST):
 
@@
     \caption{(a) C function code listing and (b) corresponding AST diagram.}
-    \label{fig:combined_ast}
+    \label{fig:ast_struc_example}
 \end{figure}
 
-\subsubsection{AST --- Types of Features}
+\subsubsection{AST - types of features}
 
-Features extracted from the AST fall into four main categories:
+Features divide into 4 main kinds according to the nature of the information extracted from the Abstract Syntax Tree (AST):
 \begin{itemize}
-    \item \textbf{Structural:} Capture the AST's complexity (e.g., depth, number of nodes, average branching factor).
-    \item \textbf{Semantic:} Reflect semantic information such as the distribution of node types or distinct root-to-leaf paths.
-    \item \textbf{Syntactic:} Describe the programming paradigm and logical complexity (e.g., number of functions, control flow units).
-    \item \textbf{Combined:} Use-case specific features that combine the aforementioned strategies for tasks like authorship attribution or 
-    function name classification.
+    \item \textbf{Structural} \label{item:structural}
+    Structural features capture the structural complexity of AST.
+These are often some numerical quantitative values like depth of the graph, number of nodes, or average branching factor of internal nodes.
+    \item \textbf{Semantic} \\
+    Semantic features reflect the semantic information encoded in AST.
+It could be as simple as the distribution of node kinds (that are defined by the compiler that creates the AST) or the distribution of distinct root-to-leaf paths.
+    \item \textbf{Syntactic} \\
+    Syntactic features describe the paradigm in which the source code is written and the logical complexity of the program.
+These might include the number of functions or functor structures, or the number of control flow units.
+    \item \textbf{Combined} \\
+    Combined features are use-case specific, and it is up to the data analyst to design the best fit solution for information extraction using the combinations and alternations of the aforementioned strategies.
+Combined features are also the ones that are usually used when tackling real-world problems like source code authorship attribution or function name classification.
 \end{itemize}
 
-\subsubsection{Halstead Complexity Measures}
+\subsubsection{Halstead complexity measures}
 
-To evaluate function complexity, we used Halstead complexity measures (Difficulty, Volume, Effort) as defined in \cite{halstead}. 
-These measures were aggregated into a single value using a weighted sum with ratios 4:3:3. (Halstead length was not incorporated, as it was analyzed separately.)
+For evaluation of complexity of examined functions we have used Halstead complexity 
+measures Difficulty, Volume, Effort \cite{halstead}. To aggregate them into single value, 
+we used weighted sum of the three metrics weighting them in ratios 4:3:3 respectively. 
+We did not incorporate Halstead length because we did separate analysis on length.
 
 \subsection{Our Data Representation}
-
-For training the models, a prediction task data pipeline was constructed as follows:
+For training the models, we constructed our prediction task data pipeline like this:
 
@@
 
-Various representations of each function’s raw source code were utilized:
-\begin{itemize}
-    \item Word-tokenized TF-IDF vectors for Random Forests.
-    \item Byte Pair Encoding (BPE) of raw source code for BERT.
-    \item Linearized AST pre-order traversal for BERT.
-    \item A set of 600 randomly selected node-to-node paths for AttentionNN.
-\end{itemize}
-The PsycheC project's C language frontend parser \cite{psychec} was used to convert code into ASTs. Models were trained in the 
-TensorFlow.Keras framework using different author set sizes (110, 27, 11, and 3) with minimum function counts per author of 25, 40, 50, and 58 respectively.
+We utilized various 
+types of representations of single function's raw source code to provide the models with data.
+These include word tokenized TF-IDF vectors for Random Forests, BPE of raw source code for
+BERT, linearized AST pre-order traversal for BERT and representation set of randomly selected 600 node-to-node paths for AttentionNN.  
+PsycheC project's C language frontend parser \cite{psychec} was used for parsing the code into ASTs.
+Each model was trained in Tensorflow.Keras framework and with different author set sizes to evaluate scalabillity.
+The different set sizes were
+110, 27, 11, 3 with each having increased demands on minimal function count than the one before. The minimal function counts per author for the sizes were 25, 40, 50, 58 respectively.
+Table \ref{tab:basic_props} summarizes the basic properties of our datasets.
 
-\subsection{Data Preprocessing \& Exploratory Data Analysis}
+\begin{table}[!b]
+    \centering
+    \caption{Basic properties of our different datasets.}
+    \label{tab:basic_props} % label placed AFTER caption
+    \begin{tabular}{|c|l|l|}
+        \hline
+        \textbf{Author set size} & \textbf{No. of functions} & \textbf{Min. function count / author} \\ 
+        \hline
+        110 & 3756 & 25 \\ 
+        \hline
+        27 & 1357 & 40 \\ 
+        \hline
+        11 & 638 & 50 \\ 
+        \hline
+        3 & 212 & 58 \\ 
+        \hline
+    \end{tabular}
+\end{table}
 
-The dataset consisted of thousands of solutions from hundreds of authors. From each solution, the function and its author were extracted. 
-The largest dataset comprised $3\,004$ C language functions from GCJ solutions (2009--2018). Preprocessing involved duplicate removal,
-elimination of authors with disproportionately many samples, and exclusion of functions that could not be parsed into an AST. Frequency 
-distributions of authors and function lengths were analyzed for further insights.
+\subsection{Data preprocessing \& Exploratory data analysis}
+The dataset consisted of thousands of solutions from hundreds of authors. From each solution we extracted the function and its author.
+The largest of the different sized datasets we used consisted of $3\,756$ C language functions' source code gathered from Google Code Jam (GCJ) solutions from years 2009 to 2018.
+The data was preprocessed exclude duplicates,
+remove outlying authors that have multiple times more samples than is the average number of samples per author to reduce the models' bias towards these names. Functions 
+that could not be parsed into an AST were also eliminated to ensure consistency in the datasets. This allows for a fair assessment of each model, 
+as they utilize different representations of function data. For further analysis, we needed to see the frequency distributions of authors.
 
@@
     \caption{Comparison of models across author attribution accuracy and training time for different author set sizes.}
-    \label{tab:comparison}
+    \label{tab:comparison_models}
     \centering
@@
 
-Figure~\ref{fig:distr_names} shows that most function lengths are between 10 and 50 tokens, 
-making single-function authorship attribution challenging when only short code snippets are available.
+From the figure \ref{fig:distr_names} we can see that the most of our dataset's 
+functions lengths lie between 10 and 50 tokens which makes our goal of authorship attribution of single function
+harder than if given a whole file consisting of multiple functions or code blocks in sense that for the prediction
+most of the time we only have from 10 to 50 tokens available to find the correct author.
 
-\section{Models Setup \& Training}
+\section{Models setup \& training}
 
-Random Forests and AttentionNN models were trained using a stratified cross-validation strategy (5 folds: 3 for training,
- 1 for validation, and 1 for testing). Because BERT training requires considerably more time, it was repeated three times, with evaluation metric variance below 1\%.
+In this section we will discuss the training strategies and setup for the examined models.
+Random Forests and AttentionNN models were trained using Stratified Cross validation strategy that ensures, that each author is evenly distributed
+across the folds. Number of folds was set to 5 and 3 folds were used for training, 1 for validation and 1 for testing.
+As training of BERT takes a lot more time than the other 2, we just repeated the traing 3 times and 
+observed variance in evaluation metrics lower than 1\%.
 
 \subsection{Random Forests}
-Random Forests were chosen as a baseline for text-based authorship attribution.
-They were trained on word-level tokens (using the regex \verb|r"\b\w+\b"|).
-A grid search was performed to optimize hyperparameters:
+
+For comparison purposes, we have chosen Random Forests as standard approach to text based authorship attribution. 
+They were trained on word-level tokens using the regex pattern \verb|r"\b\w+\b"|, which matches words separated by word boundary characters.
+Firt we used grid search to optimize these hyperparameters:
 \begin{itemize}
@@
 \end{itemize}
-The best hyperparameters were then used to train the final model.
+
+and then we trained the model using the best hyperparameters found.
 
 \subsection{BERT}
-For both BERT setups, Microsoft’s pretrained \texttt{graphcodebert-base} \cite{codebert} was used.
-The graphcodebert tokenizer uses Byte Pair Encoding, which is suitable for programming languages with rigid syntax.
-Comments were removed to ensure fairness when comparing models that utilize ASTs.
-BERT was trained on two distinct representations: tokenized raw source code and tokenized AST pre-order traversal.
+
+For both BERT training setups we used Microsoft's pretrained graphcodebert-base \cite{codebert} that already understands general programming concepts like 
+control structures, loops or functions and is able to capture differences between authors faster. We also used the graphcodebert tokenizer which uses 
+Byte pair encoding that merges frequently occurring character sequences into tokens, which is appropriate for data like source code of programming language
+that has very rigid syntax. The comments in all code samples were filtered out so the attribution task
+is 'fair' across the models that utilize ASTs.
+
+As mentioned, we have trained the BERT model on two distinct data representations. Tokenized RAW source code and tokenized AST pre-order traversal
+to see how the model reacts to these representations and how useful it may be to learn from the AST structural patterns in comparison to AttentionNN 
+that utilizes node-to-node paths.
 
 \subsection{Attention Neural Network}
-The AttentionNN model leverages AST-derived node-to-node paths (path-contexts) as proposed in \cite{code2vec}.
-Each path is represented as a triplet (start node, path of node kinds, terminal node) and encoded into vector space.
-The model then learns to weigh these contexts via an attention mechanism to distinguish authors.
 
-\subsection{Evaluation Methods}
-Evaluation metrics included validation accuracy, precision, recall, training time, and Halstead complexity measures. Deep learning models were trained with the categorical cross-entropy loss function:
+To capture all three of the structural, semantic and syntactic features in a number vector that could be somehow fed into a neural network classifier. 
+Code2vec \cite{code2vec} proposes solution design that extracts so called 'path-contexts' which are encoded into vector space using the embedding values 
+of its components.
+Path-contexts capture the node-to-node path in concise manner as it consist of triplets of (start node, the path of nodes kinds, terminal node).
+Each distinct node is encoded with the data value it carries and each distinct path is encoded as a string of nodes' kinds in between the start and terminal node.
+The model then learns to distinguish the authors utilizing the attention mechanism, giving larger weights to contexts that help to reveil the author the most.
+
+\subsection{Evaluation methods}
+For evaluation purposes metrics such as validation Accuracy, Precision, Recall, Time to train model and Halstead complexity measures were captured to explain the models
+behavior with different kinds of data. The deep-learning models were trained with the Categorical Cross-Entropy loss function guiding the optimization process,
+
 \begin{equation}
@@
 \end{equation}
+
 where:
 \begin{itemize}
-    \item $\mathcal{L}_i$ is the loss for sample $i$.
-    \item $C$ is the number of classes.
-    \item $y_{i,j}$ is the one-hot encoded ground truth for class $j$.
-    \item $\hat{y}_{i,j}$ is the predicted probability for class $j$ (obtained from a softmax function).
+    \item $\mathcal{L}_i$ is the loss for the $i$-th sample.
+    \item $C$ is the number of possible classes.
+    \item $y_{i,j}$ represents the ground truth label for class $j$. It is a one-hot encoded vector, meaning:
+    \begin{itemize}
+        \item $y_{i,j} = 1$ if the sample belongs to class $j$.
+        \item $y_{i,j} = 0$ otherwise.
     \end{itemize}
-Learning curves were analyzed to assess overfitting. The overfit ratio was calculated as:
-\begin{equation}
-    \text{Overfit Ratio} = \frac{\text{Train Accuracy}}{\text{Test Accuracy}}
-\end{equation}
+    \item $\hat{y}_{i,j}$ is the predicted probability for class $j$, obtained from a softmax function.
+    \item $\log(\hat{y}_{i,j})$ takes the logarithm of the predicted probability.
+\end{itemize}
+
+We also looked at the learning curves of the deep leaning models (BERT, AttentionNN) to assess the overfit of each one and compared the models in terms of 
+accuracy for functions of different complexity obtaining interesting results revealing their strengths and weaknesses.
 
@@
     \caption{3D bar chart comparing Accuracy, Precision, and Recall across models and author set sizes.}
-    \label{fig:3d-bars}
+    \label{fig:3d_bars}
 \end{figure*}
@@
 
-The models were evaluated for scalability as the number of functions per author increased while the distinct author set size decreased.
-Table~\ref{tab:comparison} summarizes model performance (accuracy and training time) for various author set sizes.
+As we have mentioned, the models were mainly tested for their scalability with having more functions per author while decreasing the distinct author set size.
+Table \ref{tab:comparison_models} summarizes how each model performed in terms of accuracy and the time required for training (until the validation loss stopped decreasing) across diverse datasets
+and figure \ref{fig:3d_bars} visualizes the achieved performances on Accuracy, Precision and Recall. From table \ref{tab:comparison_models} we can say that BERT 
+trained on tokens extracted from the raw source code performed the best at each of the author set sizes, especially with accuracy over 90\% on the smallest set, however with the longest required time to train the model, more than
+9.5 times longer than it took to train AttentionNN to its full potential and 36.4 times longer than the time required for Random Forests which scored just 4.8 percent less and reached similar to state-of-the art results \cite{rf1} \cite{rf2}
+What also interested us is that From \texttt{set\_size=11} to \texttt{set\_size=3}, BERT improved only by 0.86\% which is the smallest improvement across all the experiments and indicates that increasing the minimum function
+count per author from 50 to 58 had little impact on the model's ability to distinguish authors. 
+On the other hand, from being the worst, BERT fed with AST pre-order traversal scored a 40\% higher under these conditions.
+Unfortunately, we were unable to find any relevant sources against which to compare our BERT results.
+Nevertheless the following experiments conducted on other models, helped us understand advantages and disadvantages of BERT model for authorship attribution.
 
-Figure~\ref{fig:conf_matrix_bert} shows the confusion matrix for the best performing model (BERT on raw source code) for an author set size of 27.
+The confusion matrix shows \ref{fig:conf_matrix_bert}, that the model performed best on 2 most represented authors
+(see \ref{fig:distr_names}). Although if we take 8 most represented authors, only 5 of them are also 
+the most accurately classified.
+
+In terms of overfit, BERT with tokenized source codes was able to generalize the datasets in effective manner reaching 1.41, 1.24, 1.05, 1.12 for overfit ratios 
+calculated like this
+
+\begin{equation}
+    \text{Overfit Ratio} = \frac{\text{Train Accuracy}}{\text{Test Accuracy}}
+\end{equation}
+
+and only the latter two are under the limit of 1.2 - being optimally fitted.
+
+Random forests reached 1.51, 1.3, 1.24, 1.15 overfit ratios which still show a considerable amount of generalization. However AttentionNN was not that successful with generalizing 
+and with the dataset consisting of 3 distinct authors it got overfit of 1.4 and only increasing with each larger dataset, meaning insufficient data is being provided for the network so it starts memorizing particular
+patterns instead of actually generalizing above authors, which does not allow the model to perform better on unseen data affecting its validation accuracy.
+
+\subsection{Comparing models' performance depending on the complexity of function}
+
+In our task of identifying an author given a single function, a natural question arises: how did the models 
+perform on functions of different complexities? The basic indicator of function complexity is its length. So we came up with strategy, 
+how to visually compare models, even when the number of each model's misclassifications is different. We scaled the models histogram bin values by 
+an index that is obtained like this:
+
+\begin{equation}
+    \text{index}_{i} = \frac{N_{\text{w}}}{N_{i}}
+\end{equation}
+
+\noindent where:
+\begin{itemize}
+    \item $N_{\text{w}}$ is the number of misclassifications of the worst-performing model.
+    \item $N_{i}$ is the number of misclassifications made by the model \( i \).
+    \item $\text{index}_{i}$ is the scaling factor for the $i$-th model.
+\end{itemize}
+
+From the figure \ref{fig:misclass_lens} the only statement we can say is that the model AttentionNN particularly struggled with functions of length lower than 40 tokens
+but did quite well on the longer ones. This will be discussed in the following subsection more deeply \ref{subsec:comp_over_quant}.
+Other models performance is not really distinguishable from this visualization, so we performed the same index strategy with the Halstead complexity metrics.
+Measuring each misclassified function's Halstead complexity and plotting it for Random Forests and BERT trained on tokenized source code, we obtained following figure.
+
+This error distribution reveals that Random Forests struggled more with Halstead complexity increasing from the marked red point compared to the BERT 
+Specifically there are 1.59 times more misclassifications made by the Random Forest classifier than there are by BERT measured from the red point.
+However, closer we look to 0 on x axis, the more errors we see made by BERT. This would be the main difference between the deep learning and 
+machine learning method. The deep learning method utilizes self-attention that enables it to capture more complex relationships but struggles
+to categorize when limited or not complex enough data is provided, not being able to find the relationships that characterize the author of the function. 
+On the other hand, the machine learning TF-IDF method struggles with the more complex data because the representation lacks context understanding and solely 
+relies on the number of occurrences in the sample.
+
+\subsection{Attention neural network - complexity over quantity}  \label{subsec:comp_over_quant}
+
+Considering the observation in figure \ref{fig:misclass_lens}, where AttentionNN performed worst on lengths between 20 and 40 tokens,
+we assesed the model in a deeper manner. Evaluating accuracies across functions distributed into bins according to 2 metrics. 
+Number of nodes in the AST and depth of the AST of the function.
+
+This way we obtain visual representation that objectively accounts for the fact that the number of errors is different.
+In the following figure we visualized the frequencies of misclassified function lengths in tokens for Random Forests, BERT and AttentionNN scaling each 
+model with its index.
+
+From the figure we can deduce, there is an emerging trend in accuracy from bins of lower value to the bins of higher value.
+Eventhough the frequency distribution of both variables indicates more frequency on the lower values we observe a lower
+accuracy. This tells that the model is learning rather from more structural complex functions than simple non-nesting and 
+control-flow divergent ones.
 
@@
 
-We also analyzed the effect of function complexity.
-Figure~\ref{fig:misclass_lens} visualizes the frequency distribution of misclassified function lengths (in tokens) for Random Forests, BERT, and AttentionNN after scaling each model’s misclassification count by an index:
-\begin{equation}
-    \text{index}_{i} = \frac{\text{Number of misclassifications of the worst performing model}}{\text{Number of misclassifications of model } i}
-\end{equation}
-
-Figure~\ref{fig:misclass_halstead} shows the distribution of misclassified function Halstead complexity for Random Forests and BERT, revealing that while Random Forests struggle with higher complexity, BERT exhibits more misclassifications when complexity is low.
-
 \begin{figure}[!b]
@@
 \end{figure}
-
-Furthermore, we assessed the AttentionNN model’s performance based on function complexity by evaluating accuracy across bins of AST node count and AST depth.
-Figure~\ref{fig:eda_distr} presents subfigures for these two metrics, indicating that the model learns better from structurally complex functions.
-
-A 3D bar chart (Figure~\ref{fig:3d-bars}) further visualizes Accuracy, Precision, and Recall across models and author set sizes. (The code for this chart uses TikZ and pgfplots.)
 
@@
 \begin{thebibliography}{00}
-    \bibitem{b1} G.
-Eason, B.
-Noble, and I.
-N.
-Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
-Trans.
-Roy.
-Soc.
-London, vol.
-A247, pp.
-529--551, April 1955.
-    \bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
-    \bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
-    \bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
-    \bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
-    \bibitem{b6} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
-J.
-Magn.
-Japan, vol.
-2, pp.
-740--741, August 1987 [Digests 9th Annual Conf.
-Magnetics Japan, p.
-301, 1982].
-    \bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
-    \bibitem{b8} D. P. Kingma and M. Welling, ``Auto-encoding variational Bayes,'' 2013, arXiv:1312.6114. [Online]. Available: https://arxiv.org/abs/1312.6114
-    \bibitem{b9} S. Liu, ``Wi-Fi Energy Detection Testbed (12MTC),'' 2023, gitHub repository. [Online]. Available: https://github.com/liustone99/Wi-Fi-Energy-Detection-Testbed-12MTC
-    \bibitem{b10} ``Treatment episode data set: discharges (TEDS-D): concatenated, 2006 to 2009.'' U.S.
-Department of Health and Human Services, Substance Abuse and Mental Health Services Administration, Office of Applied Studies, August, 2013, DOI:10.3886/ICPSR30122.v2
-    \bibitem{b11} K. Eves and J. Valasek, ``Adaptive control for singularly perturbed systems examples,'' Code Ocean, Aug. 2023. [Online]. Available: https://codeocean.com/capsule/4989235/tree
+    \bibitem{information}
+    X.
+He, A.
+H.
+Lashkari, N.
+Vombatkere, and D.
+P.
+Sharma, "Authorship attribution methods: A comprehensive survey," \textit{Information}, vol.
+15, no.
+3, article 131, 2024.
+[Online].
+Available: https://www.mdpi.com/2078-2489/15/3/131.
+[Accessed: June 15, 2024].
+
+    \bibitem{sun_ast}
+    W.
+Sun, C.
+Fang, Y.
+Miao, Y.
+You, M.
+Yuan, Y.
+Chen, Q.
+Zhang, A.
+Guo, and X.
+Guo, "Abstract Syntax Tree for Programming Language Understanding and Representation: How Far Are We?", \textit{arXiv preprint arXiv:2305.05929}, 2023.
+[Online].
+Available: https://arxiv.org/abs/2303.01017.
+Accessed: June 15, 2024.
+
+    \bibitem{codebert}
+    D.
+Guo, S.
+Ren, S.
+Lu, Z.
+Feng, D.
+Tang, S.
+Liu, L.
+Zhou, N.
+Duan, and A.
+Svyatkovskiy, "GraphCodeBERT: Pre-training Code Representations with Data Flow," \textit{arXiv preprint arXiv:2009.08366}, 2020.
+[Online].
+Available: https://arxiv.org/abs/2009.08366.
+[Accessed: June 15, 2024].
+
+    \bibitem{code2vec}
+    U.
+Alon, M.
+Zilberstein, O.
+Levy, and E.
+Yahav, "code2vec: Learning distributed representations of code," \textit{Proc.
+ACM Program.
+Lang.}, vol.
+3, no.
+POPL, pp.
+1–29, Jan.
+2019.
+doi: 10.1145/3290353.
+[Online].
+Available: https://github.com/tech-srl/code2vec.
+
+    \bibitem{halstead}
+    K. Hariprasad and J. Valarmathi, "Software complexity analysis using halstead metrics," in \textit{Proc. 2017 Int. Conf. on Comput. Syst. and Informatics (ICEI)}, 2017, pp. 1109-1113.
+
+    \bibitem{compiler_based}
+    D.
+Guo, S.
+Ren, S.
+Lu, Z.
+Feng, D.
+Tang, S.
+Liu, L.
+Zhou, N.
+Duan, and A.
+Svyatkovskiy, "Compiler-based graph representations for deep learning models of code," in \textit{Proc.
+CC 2020}, New York, NY, USA: ACM, 2020.
+
+    \bibitem{gcj}
+    J. Petrik, "Function name classifier," 2024. [Online]. Available: \url{https://github.com/Jur1cek/gcj-dataset}. [Accessed: June 15, 2024].
+
+    \bibitem{psychec}
+    L.
+T.
+C.
+Melo, R.
+G.
+Ribeiro, B.
+C.
+F.
+Guimarães, and F.
+M.
+Q.
+Pereira, ``Type inference for C: Applications to the static analysis of incomplete programs,'' \textit{ACM Trans.
+Program.
+Lang.
+Syst.}, vol.
+42, no.
+3, Art.
+no.
+15, Nov.
+2020.
+[Online].
+Available: https://doi.org/10.1145/3421472
+
+    \bibitem{pbnn}
+    T.
+Girba, A.
+Kuhn, M.
+Seeberger, and S.
+Ducasse, ``How developers drive software evolution,'' in \textit{Eighth International Workshop on Principles of Software Evolution (IWPSE'05)}, 2005, pp.
+113--122.
+[Online].
+Available: https://doi.org/10.1109/IWPSE.2005.21
+
+    \bibitem{rf1}
+    C.
+McKnight and I.
+Goldberg, ``Style counsel: Seeing the (random) forest for the trees in adversarial code stylometry,'' in \textit{Proc.
+2018 Workshop Privacy Electron.
+Soc.}, Master's thesis, Univ.
+of Waterloo, Waterloo, ON, Canada, Oct.
+2018, pp.
+138--142.
+
+    \bibitem{rf2}
+L.
+Simko, L.
+Zettlemoyer, and T.
+Kohno, ``Recognizing and imitating programmer style: Adversaries in program authorship attribution,'' in \textit{Proceedings on Privacy Enhancing Technologies (PETS)}, vol.
+2018, no.
+1, 2018, pp.
+127--144.
+[Online].
+Available: https://doi.org/10.1515/popets-2018-0007
+
 \end{thebibliography}

### 2025-03-14 my_cst/article.tex
m

--- 2025-03-14 myIEEE/article.tex
@@
 (AttentionNN) utilizing AST-derived node-to-node paths. The evaluation is 
-conducted on a dataset of 3,004 C language functions from Google Code Jam 
+conducted on a dataset of 3,756 C language functions from Google Code Jam 
 (GCJ) spanning 2009 to 2018 \cite{gcj}. Most of the works in the field
@@
 trained on tokens extracted from the raw source code performed the best at each of the author set sizes, especially with accuracy over 90\% on the smallest set, however with the longest required time to train the model, more than
-9.5 times longer than it took to train AttentionNN to its full potential and 36.4 times longer than the time required for Random Forests which scored just 4.8 percent less and reached similar to state-of-the art results \cite{rf1} \cite{rf2}
+9.5 times longer than it took to train AttentionNN to its full potential and 36.4 times longer than the time required for Random Forests which scored just 4.8 percent less.
+Our RF experiment on dataset with 110 reached results similar to this study \cite{rf1},
+where they had 525 authors, but with 2-11 files per author and also included information from comments which we filtered out.
+In \cite{rf2}, authors reached accuracy of 84.5\% on 50 authors with and average of 42.8 files each similar to our setup for
+27 authors. They scored 14\% more than we did, using the the features presented in \cite{caliskan}.
+
 What also interested us is that From \texttt{set\_size=11} to \texttt{set\_size=3}, BERT improved only by 0.86\% which is the smallest improvement across all the experiments and indicates that increasing the minimum function
@@
 
+This way we obtain visual representation that objectively accounts for the fact that the number of errors is different.
+Figure \ref{fig:misclass_lens} shows visualized frequencies of misclassified function lengths in tokens for Random Forests, BERT and AttentionNN scaling each 
+model with its index.
+
 From the figure \ref{fig:misclass_lens} the only statement we can say is that the model AttentionNN particularly struggled with functions of length lower than 40 tokens
 but did quite well on the longer ones. This will be discussed in the following subsection more deeply \ref{subsec:comp_over_quant}.
-Other models performance is not really distinguishable from this visualization, so we performed the same index strategy with the Halstead complexity metrics.
+Other models performance is not really distinguishable from this visualization, so we performed the same index strategy with the Halstead complexity metrics in figure \ref{fig:misclass_halstead}.
 Measuring each misclassified function's Halstead complexity and plotting it for Random Forests and BERT trained on tokenized source code, we obtained following figure.
@@
 
+This model showed relatively promising results, indicating increasing dataset size and providing more complex data enhances its performance.
+Paper \cite{pbnn} presents similar model reaching 0.979\% accuracy with 40 authors but 3,021 files for each in programming language Java.
+
 Considering the observation in figure \ref{fig:misclass_lens}, where AttentionNN performed worst on lengths between 20 and 40 tokens,
@@
 
-This way we obtain visual representation that objectively accounts for the fact that the number of errors is different.
-In the following figure we visualized the frequencies of misclassified function lengths in tokens for Random Forests, BERT and AttentionNN scaling each 
-model with its index.
-
-From the figure we can deduce, there is an emerging trend in accuracy from bins of lower value to the bins of higher value.
+From the figure \ref{fig:ann_acc} we can deduce, there is an emerging trend in accuracy from bins of lower value to the bins of higher value.
 Eventhough the frequency distribution of both variables indicates more frequency on the lower values we observe a lower
-accuracy. This tells that the model is learning rather from more structural complex functions than simple non-nesting and 
+accuracy. This tells that the model is learning more about the author from structural complex functions than simple non-nesting and 
 control-flow divergent ones.
 
-\begin{figure}[!b]
+\begin{figure}[!t]
     \centering
@@
 
-\begin{figure}[!b]
+\begin{figure}[!t]
     \centering
@@
 classifier could balance accuracy and efficiency. Also, using graph-based neural networks to process ASTs more effectively might push 
-authorship attribution even further. Exploring data augmentation techniques or better regularization strategies could also help deep 
+authorship attribution even further \cite{dependence_graphs}. Exploring data augmentation techniques or better regularization strategies could also help deep 
 learning models generalize better, especially with AST-based inputs.
 
-\begin{figure}[!b]
+\begin{figure}[!t]
     \centering
@@
 
+    \bibitem{caliskan}
+    A.
+Caliskan-Islam \textit{et al.}, ``De-anonymizing programmers via code stylometry,'' in \textit{Proc.
+24th USENIX Security Symp.}, 2015, pp.
+255--270.
+[Online].
+Available: https://www.usenix.org/conference/usenixsecurity15/technical-sessions/presentation/caliskan-islam
+
+    \bibitem{dependence_graphs}
+    R.
+Chen, L.
+Hong, C.
+Chunyan Lü and W.
+Deng, ``Author Identification of Software Source Code with Program Dependence Graphs,'' in \textit{Proc.
+2010 IEEE 34th Annual Computer Software and Applications Conference Workshops}, Seoul, Korea (South), 2010, pp.
+281--286, doi: 10.1109/COMPSACW.2010.56.
+[Online].
+Available: https://doi.org/10.1109/COMPSACW.2010.56
+
 \end{thebibliography}

### 2025-06-22 my_cst/article.tex
strange format

@@
 
-% Required packages for IEEEtran
 \usepackage{cite}
@@
 \usepackage{xcolor}
-
-% Additional packages from your article (ensuring they do not interfere with IEEE margins)
 \usepackage{subcaption}
@@
 \usepackage{url}
+\usepackage{float}
 
-% Do not load geometry or bookmark in IEEEtran
-%\usepackage{geometry}
-%\usepackage{bookmark}
+\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
+    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
 
@@
 
-\title{Comparing Machine Learning and Deep Learning Models with Attention Mechanisms for Multiple Source Code Authorship Attribution\\
-\thanks{Identify applicable funding agency here. If none, delete this.}
+\title{Comparing Machine Learning and Deep Learning Models with Attention Mechanisms for Multiple Source Code Authorship Attribution
+\thanks{The work reported here was supported by the Cultural and Educational Grant Agency of Slovak Republic (KEGA) under grant No. KG 014STU-4/2024.}
 }
@@
 \author{
-\IEEEauthorblockN{Richard Čerňanský}
-\IEEEauthorblockA{\textit{Department Name}\\
-\textit{Organization}\\
-City, Country\\
-email@example.com}
+\IEEEauthorblockN{1\textsuperscript{st} Bc. Richard Čerňanský}
+\IEEEauthorblockA{
+\textit{Faculty of Informatics and Information Technologies} \\
+\textit{Slovak University of Technology}\\
+Bratislava, Slovakia \\
+richard.cernansky3@gmail.com}
 \and
-\IEEEauthorblockN{Ing. Juraj Petrík}
-\IEEEauthorblockA{\textit{Department Name}\\
-\textit{Organization}\\
-City, Country\\
-email@example.com}
+\IEEEauthorblockN{2\textsuperscript{nd} Ing. Juraj Petrík}
+\IEEEauthorblockA{
+\textit{Faculty of Informatics and Information Technologies} \\
+\textit{Slovak University of Technology}\\
+Bratislava, Slovakia \\
+juraj.petrik@stuba.sk}
 }
@@
 \begin{IEEEkeywords}
-Authorship attribution, machine learning, deep learning, BERT, attention mechanisms, source code analysis.
+    Authorship attribution, machine learning, deep learning, BERT, attention mechanisms, source code analysis
     \end{IEEEkeywords}
@@
 Despite the success of deep learning models in natural language 
-processing (NLP), their application to source code authorship attribution on different dataset sizes. 
+processing (NLP), their application to source code authorship attribution on different dataset sizes 
 remains relatively unexplored. Unlike natural language, source code 
@@
 
-This study systematically compares machine learning and deep learning 
+In the first part we present important related work results in the field of source 
+code authorship attribution. Then we continue with explaining methods and techniques 
+that we used for developing and evaluating comparative framework of different models.
+Section 3 describes the models in more detail.
+The main part of this study systematically compares machine learning and deep learning 
 models, including Random Forests, BERT trained on tokenized source code, 
@@
 conducted on a dataset of 3,756 C language functions from Google Code Jam 
-(GCJ) spanning 2009 to 2018 \cite{gcj}. Most of the works in the field
+(GCJ) spanning 2009 to 2018 \citet{petrik}. Most of the works in the field
 predict the authors based on files consisting of multiple language constructs, our 
@@
 The models are assessed based on their accuracy, precision, recall, and training efficiency across different 
-author set sizes.
-
-Through this comparison, we aim to determine which model best balances 
+author set sizes. Through this comparison, we aim to determine which model best balances 
 accuracy and computational efficiency, as well as the effectiveness of 
 different function representations for authorship attribution.
+
+\section{Related Work}
+
+\begin{table*}[!t]
+    \centering
+    \includegraphics[width=0.7\linewidth]{figures/dataset_table.png}
+    \caption{Table summarizing state-of-the-art works in Source code Authorship Attribution (AA) with machine learning approach \citet{datasets}.}
+    \label{tab:datasets_table}
+\end{table*}
+
+A number of studies have explored source code authorship attribution using varying 
+data representations and model architectures. Early work focused on using statistical
+models trained on token frequency vectors or stylometric features. More recent works, 
+displayed on the table \ref{tab:datasets_table} show that traditional Machine
+Learning (ML) models like Random Forest (RF), Support Vector Machine (SVM) or
+Neural Network (NN) can provide reliable results on different sized datasets. 
+Specifically, Caliskan et al. \citet{caliskan} used TF-IDF with RF on 
+Java code and demonstrated strong attribution capabilities using features extracted 
+from abstract syntax trees and code formatting. However, there is no common 
+agreement among researchers for selecting the best machine learning classifier 
+and data features for source code AA.
+
+In recent years, the development of transformer-based models like Bidirectional Encoder Transformer (BERT) has 
+significantly impacted Natural 
+Language Processing (NLP), particularly in areas such as Natural language 
+understanding and generation (NLG). However, their potential in the field 
+of authorship attribution, especially for source code, remains relatively 
+underexplored. In some works, they were compared to the state-of-the-art 
+models, showing promising results \citet{fi}. We are particularly interested 
+in pretrained models such as CodeBERT and GraphCodeBERT \citet{codebert}, 
+to source code. These models were initially developed for general code 
+understanding tasks (e.g., function naming, defect detection) but have shown 
+promising results in Authorship Attribution task when fine-tuned on labeled datasets. 
+Choi et. al.~\citet{choi} shows that Large Language Models (LLM) can be successfully utilized in 
+code authorship attribution tasks. However, their ultimate utility in 
+real-world deployment will struggle with costs, sophisticated prompt engineering, 
+model selection, hyperparameter
+choices, and an understanding of the models’ sensitivity to
+context.
+
+Alon et al. introduced code2vec \citet{code2vec}, an attention-based neural 
+network that learns embeddings from paths within Abstract Syntax Tree (AST). 
+This architecture enables learning from structural representations of code while 
+remaining lightweight compared to transformer models. Attention Neural 
+Network (ANN), a similar concept, leverages node-to-node ANN paths to 
+predict function authorship. 
+Bogomolov et. al~\citet{bogomolov} also examined path-based models on JAVA dataset 
+consisting of 200 000 methods and reached 
+accuracies around 98\% with Path-based Neural Network and Path-based Random Forests.
+
+However, much of the existing work focuses on large datasets or file-level 
+attribution. For example, studies using GitHub datasets typically assume 
+access to full repositories or complete files. In contrast, authorship 
+attribution at the function level, especially on small datasets such as 
+our collection of GCJ solutions, remains unexplored. Moreover, another 
+issue lies in the absence of standardized benchmark datasets for the task, 
+as the current research relies on datasets with different properties in terms 
+of size, author set size, source code background and other.
+
+Moreover, little attention has been paid to the impact of function complexity 
+on model performance. Most prior evaluations rely on global accuracy metrics 
+without analyzing how model effectiveness varies across functions of differing 
+structural or semantic richness.
 
@@
 
-In the paper \cite{information}, authors present division of features based on the method they use for extraction:
+In the paper \citet{information}, authors present division of features based on the method they use for extraction:
 
@@
 
-\subsection{Abstract Syntax Tree (AST)}
+\begin{figure}[!b]
 
-\quad Based on the full definition \cite{sun_ast} Abstract syntax tree (AST) is a tree-based representation of source code, where nodes represent various elements of the source code 
-and paths represent relationships of the structure. AST is an inseparable part of compilation process for many reasons. It is used as an 
-intermediate representation of program utilized for optimization and generation of machine code. Nodes of the tree can be either internal 
-(non-leaf) or leafs. Internal nodes define the program's constructs or operations for their children. Leaf nodes store the actual textual value.
-
-In the figure \ref{fig:ast_struc_example} is an example of a simple C function and its corresponding Abstract Syntax Tree (AST):
-
-\begin{figure}[!t]
     \centering
@@
     \begin{minipage}{\columnwidth}
-    \begin{lstlisting}[caption={Simple C function for addition}, label={lst:add}, basicstyle=\footnotesize\ttfamily, frame=single, numbers=left]
+        \begin{lstlisting}[
+            caption={Simple C function for addition},
+            label={lst:add},
+            basicstyle=\footnotesize\ttfamily,
+            frame=single,
+            numbers=left,
+            numberstyle=\tiny\color{gray},
+            numbersep=5pt,         % distance from code to line number
+            xleftmargin=2em        % indent the whole listing 2em from the left
+          ]
           int add(int a, int b) {
@@
     }
-    \caption{(a) C function code listing and (b) corresponding AST diagram.}
+        % Figure 3: AST example
+    \caption{(a) C function for addition; (b) its Abstract Syntax Tree.}
+    \Description{Left panel: numbered C code listing “int add(int a, int b) { return a + b; }”.
+Right panel: AST tree with nodes for TranslationUnit, FunctionDefinition, BasicTypeSpecifier “int”, FunctionDeclarator “add(…)”, CompoundStatement, ReturnStatement, AddExpression, and IdentifierName leaves “a” and “b”.}
     \label{fig:ast_struc_example}
 \end{figure}
+
+\subsection{Abstract Syntax Tree (AST)}
+
+\quad Based on the full definition \citet{sun_ast} Abstract syntax tree (AST) is a tree-based representation of source code, where nodes represent various elements of the source code 
+and paths represent relationships of the structure. AST is an inseparable part of compilation process for many reasons. It is used as an 
+intermediate representation of program utilized for optimization and generation of machine code. Nodes of the tree can be either internal 
+non-leaf or leaf nodes. Internal nodes define the program's constructs or operations for their children. Leaf nodes store the actual textual value.
+
+In the figure \ref{fig:ast_struc_example} is an example of a simple C function and its corresponding Abstract Syntax Tree (AST):
 
@@
 
-For evaluation of complexity of examined functions we have used Halstead complexity 
-measures Difficulty, Volume, Effort \cite{halstead}. To aggregate them into single value, 
-we used weighted sum of the three metrics weighting them in ratios 4:3:3 respectively. 
-We did not incorporate Halstead length because we did separate analysis on length.
+From Figure~\ref{fig:distr_lens}, we observe that functions 
+vary both in their length and the same applies also to the structural complexity of their ASTs. 
+Therefore, we decided to evaluate the models' performance in terms of function 
+complexity. There are various measures of complexity and for our purposes, 
+we selected two: the simple length in tokens and a measure called Halstead 
+complexity, which quantifies Halstead Difficulty, Volume, and Effort \citet{halstead}. 
+To aggregate them into single value, we used weighted sum of the three metrics 
+weighting them in ratios 4:3:3 respectively. We chose these ratios because it reflects both the theoretical importance 
+and the empirical scale of the three metrics. Halstead Difficulty is most directly 
+tied to cognitive complexity, so we give it a slightly larger share (40 \%), 
+while Volume (30 \%) and Effort (30 \%) each make substantial but secondary 
+contributions. Before applying these weights we normalize each metric 
+to the same range, so that the weight split yields a balanced 
+composite score—capturing “how hard” (Difficulty), “how big” (Volume), 
+and “how much work” (Effort) in one single complexity value. We did not 
+incorporate Halstead length because we did separate analysis on length.
 
@@
 For training the models, we constructed our prediction task data pipeline like this:
-
-\begin{figure}[!t]
-    \centering
-    \includegraphics[width=3.5cm,height=5cm]{figures/high_level_prediciton_task.png}
-    \caption{\cite{compiler_based} High-level overview of the function classification prediction task.}
-    \label{fig:predictive_task}
-\end{figure}
 
@@
 BERT, linearized AST pre-order traversal for BERT and representation set of randomly selected 600 node-to-node paths for AttentionNN.  
-PsycheC project's C language frontend parser \cite{psychec} was used for parsing the code into ASTs.
-Each model was trained in Tensorflow.Keras framework and with different author set sizes to evaluate scalabillity.
+PsycheC project's C language frontend parser \citet{psychec} was used for parsing the code into ASTs.
+Each model was trained in Tensorflow.Keras framework and with different author set sizes to evaluate scalability.
 The different set sizes were
@@
     \centering
-    \caption{Basic properties of our different datasets.}
+    \caption{Basic properties of our various datasets.}
     \label{tab:basic_props} % label placed AFTER caption
@@
 
+\begin{figure}[!b]
+    \centering
+    \includegraphics[width=3.5cm,height=5cm]{figures/high_level_prediciton_task.png}
+    \caption{\citet{compiler_based} Pipeline overview for source-code authorship classification.}
+    \Description{Flowchart showing raw functions parsed into tokens and ASTs, features extracted (TF-IDF, AST paths, BERT embeddings), and three classifiers (Random Forest, BERT, AttentionNN) producing author predictions.}
+
+    \label{fig:predictive_task}
+\end{figure}
+
 \subsection{Data preprocessing \& Exploratory data analysis}
@@
 
-\begin{table*}[!t]
+\begin{figure}[!t]
+    \centering
+    \begin{subfigure}{0.48\textwidth}
+      \includegraphics[width=\linewidth]{figures/distribution_names.pdf}
+      \caption{Functions per author.}
+      \label{fig:distr_names}
+    \end{subfigure}%
+    \hfill
+    \begin{subfigure}{0.48\textwidth}
+      \includegraphics[width=\linewidth]{figures/distr_lens_tokens.pdf}
+      \caption{Function length in tokens.}
+      \label{fig:distr_lens}
+    \end{subfigure}
+    \caption{Frequency distributions of functions per author and of function lengths.}
+    \Description{Two histograms: left shows the long‐tail distribution of functions across authors; right shows most functions fall between 10 and 50 tokens.}
+    \label{fig:eda_distr}
+  \end{figure}
+
+From the figure \ref{fig:distr_lens} we can see that the most of our dataset's 
+functions lengths lie between 10 and 50 tokens which makes our goal of authorship attribution of single function
+harder than if given a whole file consisting of multiple functions or code blocks in sense that for the prediction
+most of the time we only have from 10 to 50 tokens available to find the correct author.
+After having all of the functions' source codes ready in the suitable form for the model to train we 
+performed the training and evaluation on 4 different model setups that are presented in next section.
+The high-level overview of the pipeline is illustrated in the figure~\ref{fig:predictive_task}.
+
+\section{Models setup \& training}
+
+In this section we will discuss the training strategies and setup for the examined models.
+Random Forests and AttentionNN models were trained using Stratified Cross validation strategy that ensures, that each author is evenly distributed
+across the folds. Number of folds was set to 5 and 3 folds were used for training, 1 for validation and 1 for testing.
+As training of BERT takes a lot more time than the other 2, we just repeated the training 3 times and 
+observed variance in evaluation metrics lower than 1\%.
+
+\subsection{Random Forests}
+
+For comparison purposes, we have chosen Random Forests as standard approach to text based authorship attribution. 
+They were trained on word-level tokens using the regex pattern \verb|r"\b\w+\b"|, which matches words separated by word boundary characters.
+First we used Grid search to optimize these hyperparameters:
+\begin{itemize}
+    \item \textbf{n\_estimators}: \{50, 100, 200, 500\}
+    \item \textbf{max\_depth}: \{10, 20, 30, None\}
+    \item \textbf{min\_samples\_split}: \{2, 5, 10\}
+    \item \textbf{min\_samples\_leaf}: \{1, 2, 4\}
+    \item \textbf{max\_features}: \{"sqrt", "log2", None\}
+    \item \textbf{bootstrap}: \{True, False\}
+\end{itemize}
+
+and then we trained the model using the best hyperparameters found. Since each 
+dataset was tuned separately via grid search, the optimal hyperparameters vary 
+and are not listed here.
+
+\subsection{Bidirectional Encoder Representations from Transformers - BERT}
+
+Transformer-based models such as BERT leverage the self-attention 
+mechanism to compute contextual embeddings for each code token by weighting 
+interactions across the entire sequence, thus capturing both local syntactic 
+patterns and long-range semantic dependencies. Pre-trained on massive corpora 
+via masked language modeling and next-sentence prediction, BERT encodes code 
+snippets into rich representations summarized by the special \texttt{[CLS]} token. 
+By fine-tuning this encoder with a lightweight classification head on labeled 
+code–author pairs, the model learns author-specific stylistic cues like naming conventions, formatting habits, 
+and preferred idioms.
+
+For both BERT training setups we used Microsoft's pre-trained graphcodebert-base \citet{codebert} that already understands general programming concepts like 
+control structures, loops or functions and is able to capture differences between authors faster. We also used the graphcodebert tokenizer which uses 
+Byte pair encoding that merges frequently occurring character sequences into tokens, which is appropriate for data like source code of programming language
+that has very rigid syntax. The comments in all code samples were filtered out so the attribution task
+is 'fair' across the models that utilize ASTs.
+
+As mentioned, we have trained the BERT model on two distinct data representations. Tokenized RAW source code and tokenized AST pre-order traversal
+to see how the model reacts to these representations and how useful it may be to learn from the AST structural patterns in comparison to AttentionNN 
+that utilizes node-to-node paths.
+
+\begin{table*}[!ht]
     \renewcommand{\arraystretch}{1.3}
@@
 
-\begin{figure}[!t]
-    \centering
-    \begin{subfigure}{0.48\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{figures/distribution_names.pdf}
-        \caption{Frequency distribution of author names}
-        \label{fig:distr_names}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.48\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{figures/distr_lens_tokens.pdf}
-        \caption{Frequency distribution of function lengths}
-        \label{fig:disrt_lens}
-    \end{subfigure}
-    \caption{Important frequency distributions of our dataset.}
-    \label{fig:eda_distr}
-\end{figure}
-
-From the figure \ref{fig:distr_names} we can see that the most of our dataset's 
-functions lengths lie between 10 and 50 tokens which makes our goal of authorship attribution of single function
-harder than if given a whole file consisting of multiple functions or code blocks in sense that for the prediction
-most of the time we only have from 10 to 50 tokens available to find the correct author.
-
-\section{Models setup \& training}
-
-In this section we will discuss the training strategies and setup for the examined models.
-Random Forests and AttentionNN models were trained using Stratified Cross validation strategy that ensures, that each author is evenly distributed
-across the folds. Number of folds was set to 5 and 3 folds were used for training, 1 for validation and 1 for testing.
-As training of BERT takes a lot more time than the other 2, we just repeated the traing 3 times and 
-observed variance in evaluation metrics lower than 1\%.
-
-\subsection{Random Forests}
-
-For comparison purposes, we have chosen Random Forests as standard approach to text based authorship attribution. 
-They were trained on word-level tokens using the regex pattern \verb|r"\b\w+\b"|, which matches words separated by word boundary characters.
-Firt we used grid search to optimize these hyperparameters:
-\begin{itemize}
-    \item \textbf{n\_estimators}: \{50, 100, 200, 500\}
-    \item \textbf{max\_depth}: \{10, 20, 30, None\}
-    \item \textbf{min\_samples\_split}: \{2, 5, 10\}
-    \item \textbf{min\_samples\_leaf}: \{1, 2, 4\}
-    \item \textbf{max\_features}: \{"sqrt", "log2", None\}
-    \item \textbf{bootstrap}: \{True, False\}
-\end{itemize}
-
-and then we trained the model using the best hyperparameters found.
-
-\subsection{BERT}
-
-For both BERT training setups we used Microsoft's pretrained graphcodebert-base \cite{codebert} that already understands general programming concepts like 
-control structures, loops or functions and is able to capture differences between authors faster. We also used the graphcodebert tokenizer which uses 
-Byte pair encoding that merges frequently occurring character sequences into tokens, which is appropriate for data like source code of programming language
-that has very rigid syntax. The comments in all code samples were filtered out so the attribution task
-is 'fair' across the models that utilize ASTs.
-
-As mentioned, we have trained the BERT model on two distinct data representations. Tokenized RAW source code and tokenized AST pre-order traversal
-to see how the model reacts to these representations and how useful it may be to learn from the AST structural patterns in comparison to AttentionNN 
-that utilizes node-to-node paths.
-
 \subsection{Attention Neural Network}
@@
 To capture all three of the structural, semantic and syntactic features in a number vector that could be somehow fed into a neural network classifier. 
-Code2vec \cite{code2vec} proposes solution design that extracts so called 'path-contexts' which are encoded into vector space using the embedding values 
+Code2vec \citet{code2vec} proposes solution design that extracts so called 'path-contexts' which are encoded into vector space using the embedding values 
 of its components.
@@
 Each distinct node is encoded with the data value it carries and each distinct path is encoded as a string of nodes' kinds in between the start and terminal node.
-The model then learns to distinguish the authors utilizing the attention mechanism, giving larger weights to contexts that help to reveil the author the most.
+The model then learns to distinguish the authors utilizing the attention mechanism, giving larger weights to contexts that help to reveal the author the most.
 
@@
     \end{tikzpicture}
-    \caption{3D bar chart comparing Accuracy, Precision, and Recall across models and author set sizes.}
+    \caption{Accuracy, precision, and recall across models and author set sizes.}
+\Description{3D bar plot where the x-axis lists models (RF, BERT raw, BERT AST, AttentionNN), the y-axis shows author set sizes (110, 27, 11, 3), and the z-axis gives metric values; bars are blue for accuracy, red for precision, and brown for recall.}
     \label{fig:3d_bars}
@@
 9.5 times longer than it took to train AttentionNN to its full potential and 36.4 times longer than the time required for Random Forests which scored just 4.8 percent less.
-Our RF experiment on dataset with 110 reached results similar to this study \cite{rf1},
+Our RF experiment on dataset with 110 reached results similar to this study \citet{rf1},
 where they had 525 authors, but with 2-11 files per author and also included information from comments which we filtered out.
-In \cite{rf2}, authors reached accuracy of 84.5\% on 50 authors with and average of 42.8 files each similar to our setup for
-27 authors. They scored 14\% more than we did, using the the features presented in \cite{caliskan}.
+In \citet{rf2}, authors reached accuracy of 84.5\% on 50 authors with and average of 42.8 files each similar to our setup for
+27 authors. They scored 14\% more than we did, using the the features presented in \citet{caliskan}.
 
-What also interested us is that From \texttt{set\_size=11} to \texttt{set\_size=3}, BERT improved only by 0.86\% which is the smallest improvement across all the experiments and indicates that increasing the minimum function
+What also interested us is that from \texttt{set\_size=11} to \texttt{set\_size=3}, BERT improved only by 0.86\% which is the smallest improvement across all the experiments and indicates that increasing the minimum function
 count per author from 50 to 58 had little impact on the model's ability to distinguish authors. 
@@
 (see \ref{fig:distr_names}). Although if we take 8 most represented authors, only 5 of them are also 
-the most accurately classified.
+the most accurately classified. Additionally, BERT and other models were not biased towards the most represented authors when tested on the 
+less represented authors. The only type of bias we have seen was on author that was not in the most represented group. Therefore, we conclude 
+that the bias arises not from class imbalance but from the characteristics of each author’s 
+functions (e.g., their purpose and functionality).
 
@@
 
+\begin{figure}[H]
+    \centering
+    \hspace*{-25pt}\includegraphics[width=1.2\columnwidth]{figures/misclass_lens.pdf}
+    % Figure 8: Misclassifications by length
+\caption{Misclassification counts by function length for each model.}
+\Description{Overlaid line plots (blue=RF, red=BERT, brown=AttentionNN) showing number of misclassified functions binned by token length.}
+    \label{fig:misclass_lens}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \hspace*{-25pt}\includegraphics[width=1.2\columnwidth]{figures/misclass_hs.png}
+    % Figure 9: Misclassifications by complexity
+\caption{Misclassification counts by Halstead complexity for RF and BERT.}
+\Description{Overlaid histograms (blue=RF, red=BERT) of misclassified functions versus composite Halstead complexity scores.}
+    \label{fig:misclass_halstead}
+\end{figure}
+
 \subsection{Attention neural network - complexity over quantity}  \label{subsec:comp_over_quant}
@@
 This model showed relatively promising results, indicating increasing dataset size and providing more complex data enhances its performance.
-Paper \cite{pbnn} presents similar model reaching 0.979\% accuracy with 40 authors but 3,021 files for each in programming language Java.
+Paper \citet{pbnn} presents similar model reaching 0.979\% accuracy with 40 authors but 3,021 files for each in programming language Java.
 
 Considering the observation in figure \ref{fig:misclass_lens}, where AttentionNN performed worst on lengths between 20 and 40 tokens,
-we assesed the model in a deeper manner. Evaluating accuracies across functions distributed into bins according to 2 metrics. 
+we assessed the model in a deeper manner. Evaluating accuracies across functions distributed into bins according to 2 metrics. 
 Number of nodes in the AST and depth of the AST of the function.
 
-From the figure \ref{fig:ann_acc} we can deduce, there is an emerging trend in accuracy from bins of lower value to the bins of higher value.
-Eventhough the frequency distribution of both variables indicates more frequency on the lower values we observe a lower
-accuracy. This tells that the model is learning more about the author from structural complex functions than simple non-nesting and 
-control-flow divergent ones.
-
-\begin{figure}[!t]
-    \centering
-    \hspace*{-25pt}\includegraphics[width=1.2\columnwidth]{figures/misclass_lens.pdf}
-    \caption{Frequency distribution of misclassified function lengths (in tokens) for Random Forests, BERT, and AttentionNN.}
-    \label{fig:misclass_lens}
-\end{figure}
-
-\begin{figure}[!t]
-    \centering
-    \hspace*{-25pt}\includegraphics[width=1.2\columnwidth]{figures/misclass_hs.png}
-    \caption{Frequency distribution of misclassified function Halstead complexity for Random Forests and BERT.}
-    \label{fig:misclass_halstead}
-\end{figure}
-
-\begin{figure}[!t]
-    \centering
-    % Scale to the column width (100%)
-    \includegraphics[width=\columnwidth]{figures/conf_matrix_sourcode_bert_ts27.pdf}
-    \caption{Confusion matrix for BERT trained on raw source code (author set size 27).}
-    \label{fig:conf_matrix_bert}
-\end{figure}
-
-\section{Discussion}
-
-Our findings show that deep learning models, especially BERT, perform better than traditional machine learning methods for authorship attribution 
-of single functions. But this comes at a cost—BERT needs way more computation and training time. On the other hand, Random Forest with TF-IDF 
-tokenization held up surprisingly well, especially with larger author sets, and trained very fast.
-
-A key takeaway is that BERT trained on raw source code tokens consistently had the best accuracy across all dataset sizes. However, 
-its improvement from \texttt{set\_size=11} to \texttt{set\_size=3} was only 0.86\%, meaning increasing the minimum function count per author 
-after a certain point doesn’t do much. Interestingly, BERT using AST pre-order traversal struggled with large author sets but got way better 
-when we reduced the dataset to three authors. This shows how important it is to choose the right input representation for deep learning models.
-
-Looking at misclassifications, each model had its weak points. BERT had trouble with short functions that had low Halstead complexity— they 
-don’t provide enough meaningful context to determine the author. Random Forests, which rely more on statistical patterns than contextual 
-understanding, started failing as function complexity increased. This makes sense since TF-IDF representations capture token frequency but 
-don’t really understand code structure.
-
-AttentionNN, which works with AST-derived node-to-node paths, preferred structurally complex functions over simple ones. This means that 
-networks using structural information may do better with more complex functions rather than short, flat ones. However, AttentionNN didn’t 
-generalize well when data was limited, and it started overfitting when trained on small author sets. This shows we either need bigger 
-datasets or more regularization to prevent overfitting.
-
-The results highlight a clear trade-off between model complexity, training time, and accuracy. Deep learning models, while powerful, 
-require serious computational resources, making them harder to scale in real-world use cases. Random Forests might not be as accurate, 
-but they’re super efficient and a solid alternative when speed is a concern.
-
-Going forward, it would be interesting to combine these methods. For example a hybrid approach using BERT embeddings with Random Forest 
-classifier could balance accuracy and efficiency. Also, using graph-based neural networks to process ASTs more effectively might push 
-authorship attribution even further \cite{dependence_graphs}. Exploring data augmentation techniques or better regularization strategies could also help deep 
-learning models generalize better, especially with AST-based inputs.
-
-\begin{figure}[!t]
+\begin{figure}[!b]
     \centering
@@
     \end{subfigure}
-    \caption{AttentionNN accuracy for functions with different structural complexities.}
+    \caption{AttentionNN accuracy as a function of AST complexity.}
+    \Description{Two line plots: (a) accuracy versus number of AST nodes, showing higher accuracy on larger trees; (b) accuracy versus AST depth, showing improvement on deeper structures.}
     \label{fig:ann_acc}
+\end{figure}
+
+From the figure \ref{fig:ann_acc} we can deduce, there is an emerging trend in accuracy from bins of lower value to the bins of higher value.
+Even though the frequency distribution of both variables indicates more frequency on the lower values we observe a lower
+accuracy. This tells that the model is learning more about the author from structural complex functions than simple non-nesting and 
+control-flow divergent ones.
+
+\section{Discussion}
+
+Our findings show that deep learning models, especially BERT, perform better than traditional machine learning methods for authorship attribution 
+of single functions. But this comes at a cost—BERT needs way more computation and training time. On the other hand, Random Forest with TF-IDF 
+tokenization held up surprisingly well, especially with larger author sets, and trained very fast.
+
+A key takeaway is that BERT trained on raw source code tokens consistently had the best accuracy across all dataset sizes. However, 
+its improvement from \texttt{set\_size=11} to \texttt{set\_size=3} was only 0.86\%, meaning increasing the minimum function count per author 
+after a certain point doesn’t do much. Interestingly, BERT using AST pre-order traversal struggled with large author sets but got way better 
+when we reduced the dataset to three authors. This shows how important it is to choose the right input representation for deep learning models.
+
+Looking at misclassifications, each model had its weak points. BERT had trouble with short functions that had low Halstead complexity— they 
+do not provide enough meaningful context to determine the author. Random Forests, which rely more on statistical patterns than contextual 
+understanding, started failing as function complexity increased. This makes sense since TF-IDF representations capture token frequency but 
+do not really understand code structure.
+
+AttentionNN, which works with AST-derived node-to-node paths, preferred structurally complex functions over simple ones. This means that 
+networks using structural information may do better with more complex functions rather than short, flat ones. However, AttentionNN did not 
+generalize well when data was limited, and it started overfitting when trained on small author sets. This shows we either need bigger 
+datasets or more regularization to prevent overfitting.
+
+The results highlight a clear trade-off between model complexity, training time, and accuracy. Deep learning models, while powerful, 
+require serious computational resources, making them harder to scale in real-world use cases. Random Forests might not be as accurate, 
+but they are super efficient and a solid alternative when speed is a concern.
+
+Going forward, it would be interesting to combine these methods. For example a hybrid approach using BERT embeddings with Random Forest 
+classifier could balance accuracy and efficiency. Also, using graph-based neural networks to process ASTs more effectively might push 
+authorship attribution even further \citet{dependence_graphs}. Exploring data augmentation techniques or better regularization strategies could also help deep 
+learning models generalize better, especially with AST-based inputs.
+
+\begin{figure}[!t]
+    \centering
+    % Scale to the column width (100%)
+    \includegraphics[width=\columnwidth]{figures/conf_matrix_sourcode_bert_ts27.pdf}
+    \caption{Confusion matrix for BERT trained on raw source code (27 authors).}
+\Description{27×27 heatmap with true authors on rows and predicted on columns; dark diagonal cells show high correct predictions, especially for the two most frequent authors, lighter off-diagonals indicate errors.}
+    \label{fig:conf_matrix_bert}
 \end{figure}
@@
 
-\begin{thebibliography}{00}
-    \bibitem{information}
-    X.
-He, A.
-H.
-Lashkari, N.
-Vombatkere, and D.
-P.
-Sharma, "Authorship attribution methods: A comprehensive survey," \textit{Information}, vol.
-15, no.
-3, article 131, 2024.
-[Online].
-Available: https://www.mdpi.com/2078-2489/15/3/131.
-[Accessed: June 15, 2024].
+\begin{acks}
+    The work reported here was supported by the 
+    Cultural and Educational Grant Agency of Slovak 
+    Republic (KEGA) under grant No. KG 014STU-4/2024.
+\end{acks}
 
-    \bibitem{sun_ast}
-    W.
-Sun, C.
-Fang, Y.
-Miao, Y.
-You, M.
-Yuan, Y.
-Chen, Q.
-Zhang, A.
-Guo, and X.
-Guo, "Abstract Syntax Tree for Programming Language Understanding and Representation: How Far Are We?", \textit{arXiv preprint arXiv:2305.05929}, 2023.
-[Online].
-Available: https://arxiv.org/abs/2303.01017.
-Accessed: June 15, 2024.
-
-    \bibitem{codebert}
-    D.
-Guo, S.
-Ren, S.
-Lu, Z.
-Feng, D.
-Tang, S.
-Liu, L.
-Zhou, N.
-Duan, and A.
-Svyatkovskiy, "GraphCodeBERT: Pre-training Code Representations with Data Flow," \textit{arXiv preprint arXiv:2009.08366}, 2020.
-[Online].
-Available: https://arxiv.org/abs/2009.08366.
-[Accessed: June 15, 2024].
-
-    \bibitem{code2vec}
-    U.
-Alon, M.
-Zilberstein, O.
-Levy, and E.
-Yahav, "code2vec: Learning distributed representations of code," \textit{Proc.
-ACM Program.
-Lang.}, vol.
-3, no.
-POPL, pp.
-1–29, Jan.
-2019.
-doi: 10.1145/3290353.
-[Online].
-Available: https://github.com/tech-srl/code2vec.
-
-    \bibitem{halstead}
-    K. Hariprasad and J. Valarmathi, "Software complexity analysis using halstead metrics," in \textit{Proc. 2017 Int. Conf. on Comput. Syst. and Informatics (ICEI)}, 2017, pp. 1109-1113.
-
-    \bibitem{compiler_based}
-    D.
-Guo, S.
-Ren, S.
-Lu, Z.
-Feng, D.
-Tang, S.
-Liu, L.
-Zhou, N.
-Duan, and A.
-Svyatkovskiy, "Compiler-based graph representations for deep learning models of code," in \textit{Proc.
-CC 2020}, New York, NY, USA: ACM, 2020.
-
-    \bibitem{gcj}
-    J. Petrik, "Function name classifier," 2024. [Online]. Available: \url{https://github.com/Jur1cek/gcj-dataset}. [Accessed: June 15, 2024].
-
-    \bibitem{psychec}
-    L.
-T.
-C.
-Melo, R.
-G.
-Ribeiro, B.
-C.
-F.
-Guimarães, and F.
-M.
-Q.
-Pereira, ``Type inference for C: Applications to the static analysis of incomplete programs,'' \textit{ACM Trans.
-Program.
-Lang.
-Syst.}, vol.
-42, no.
-3, Art.
-no.
-15, Nov.
-2020.
-[Online].
-Available: https://doi.org/10.1145/3421472
-
-    \bibitem{pbnn}
-    T.
-Girba, A.
-Kuhn, M.
-Seeberger, and S.
-Ducasse, ``How developers drive software evolution,'' in \textit{Eighth International Workshop on Principles of Software Evolution (IWPSE'05)}, 2005, pp.
-113--122.
-[Online].
-Available: https://doi.org/10.1109/IWPSE.2005.21
-
-    \bibitem{rf1}
-    C.
-McKnight and I.
-Goldberg, ``Style counsel: Seeing the (random) forest for the trees in adversarial code stylometry,'' in \textit{Proc.
-2018 Workshop Privacy Electron.
-Soc.}, Master's thesis, Univ.
-of Waterloo, Waterloo, ON, Canada, Oct.
-2018, pp.
-138--142.
-
-    \bibitem{rf2}
-    L.
-Simko, L.
-Zettlemoyer, and T.
-Kohno, ``Recognizing and imitating programmer style: Adversaries in program authorship attribution,'' in \textit{Proceedings on Privacy Enhancing Technologies (PETS)}, vol.
-2018, no.
-1, 2018, pp.
-127--144.
-[Online].
-Available: https://doi.org/10.1515/popets-2018-0007
-
-    \bibitem{caliskan}
-    A.
-Caliskan-Islam \textit{et al.}, ``De-anonymizing programmers via code stylometry,'' in \textit{Proc.
-24th USENIX Security Symp.}, 2015, pp.
-255--270.
-[Online].
-Available: https://www.usenix.org/conference/usenixsecurity15/technical-sessions/presentation/caliskan-islam
-
-    \bibitem{dependence_graphs}
-    R.
-Chen, L.
-Hong, C.
-Chunyan Lü and W.
-Deng, ``Author Identification of Software Source Code with Program Dependence Graphs,'' in \textit{Proc.
-2010 IEEE 34th Annual Computer Software and Applications Conference Workshops}, Seoul, Korea (South), 2010, pp.
-281--286, doi: 10.1109/COMPSACW.2010.56.
-[Online].
-Available: https://doi.org/10.1109/COMPSACW.2010.56
-
-\end{thebibliography}
+\bibliographystyle{IEEEtran}
+\bibliography{main}
 
